\documentclass[12pt,a4paper]{article}
\usepackage{amsmath,amsfonts,amscd,amsthm}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    citecolor=black,
    urlcolor=blue
    }
\usepackage{enumitem}
%\usepackage{mathptmx}
\usepackage{dsfont}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{multirow,booktabs}
\usepackage{makecell}
\usepackage{graphics,subfig}
\usepackage{setspace}
\usepackage{changepage}
\usepackage{multicol}
\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage{geometry}
\usepackage{dcolumn}
\usepackage{cleveref}
\usepackage[most]{tcolorbox}
\tcbuselibrary{theorems}
\newlength{\tabcont}
%\setlength{\parindent}{0.0in}
%\setlength{\parskip}{0.05in}
\usepackage{natbib}
\geometry{top=2cm,bottom=2cm,left=3cm,right=2cm, headsep=0in}
%\parindent 0in
%\parskip 12pt
\onehalfspacing

% The theoremstyle for a theorem goes right above the \newtheorem command
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[section]
\newtcbtheorem[number within=section]{remark}{Remark}{
    breakable,
    colback=gray!5,
    colframe=gray!80,
    boxrule=0.8pt,
    arc=2pt,
    left=6pt,
    right=6pt,
    top=6pt,
    bottom=6pt,
    fonttitle=\bfseries
}{rem}
\tcbset{
  enhanced,
  breakable
}
\newtheorem{assumption}{Assumption}[section]
\renewcommand{\baselinestretch}{1.5} 

\title{Master Thesis Topic}
\author{Fabian Christian Schmidt}
\date{\today}

\begin{document}

\begin{titlepage}
    \begin{center}
        \vspace*{0.5cm}
            
        \huge
        \textbf{On Functional Garch Models and Their Application to Intraday Value-At-Risk}
            
        \vspace{1cm}
        \large
        Master Thesis Presented to the\\
        Department of Economics at the\\
        Rheinische Friedrich-Wilhelm-Universität Bonn
            
        \vspace{1.5cm}
            
        In Partial Fulfillment of the Requirements of the Degree of\\
        Master of Science (M.Sc.)
            
        \vspace{4cm}
            
        \large
        Supervisor: Prof. Dr. Dominik Liebl

        \vspace{1cm}
        Submitted in March 2026 by:\\
        Fabian Christian Schmidt\\
        Matriculation number: 50092796
            
    \end{center}
\end{titlepage}
\pagestyle{empty}
\tableofcontents
\clearpage
\pagestyle{plain}
\setcounter{page}{1}
\section{Introduction}
\label{sec:Introduction}
%\begin{itemize}
%    \item Start with explanation of the history of interest in risk measures
%    \item transition to GARCH models and their usage when estimating risk $\longrightarrow$ name most important papers regarding GARCH models
%    \item transition to Functional Data Analysis and then Functional GARCH and its applications (literature)
%    \item explain structure and goal of the thesis
%\end{itemize}
Interest in the quantitative assessment of financial risk has a long history, with key foundations emerging in the early $20^{\text{th}}$ century. An early milestone was Louis Bachelier’s dissertation, which introduced probabilistic modelling into finance and laid the groundwork for modern mathematical finance. More broadly, risk management aims to identify, measure, and evaluate potential losses in order to support decision-making under uncertainty. To this end, risk managers employ specific risk measures to assess the exposure associated with a particular decision or investment.\\
One of the most widely used measures of market risk, defined as the risk of losses arising from on- and off-balance-sheet positions, is value-at-risk (VaR). Early contributions to the measurement of downside risk can be found in \citet{Leavens1945}, while related concepts appear in the safety-first criterion of \citet{Roy1952} and the mean--variance framework of \citet{Markowitz1952}. However, these works did not define VaR in its modern form. VaR emerged as an operational tool for market risk measurement in the early 1990s, when J.P. Morgan introduced it internally as a standardized approach to measuring firm-wide market risk. The public release of the \textit{RiskMetrics} methodology in 1994 contributed to the widespread adoption of VaR by financial institutions. VaR was subsequently incorporated into financial regulation through the 1996 Market Risk Amendment to the Basel I Accord, which allowed banks to use internal VaR models for the calculation of regulatory capital. Since the forecasting of VaR requires an assessment of the conditional distribution of returns, models for time-varying volatility play a central role in its practical implementation.\\
In practical applications, one common approach to forecasting VaR is to model the conditional volatility of financial returns using autoregressive conditional heteroskedasticity models. The GARCH model, introduced by \citet{Bollerslev1986} as a generalization of the ARCH model of \citet{Engle1982}, differs fundamentally from standard time series models such as ARMA or VAR models. Rather than focusing on the conditional mean, ARCH and GARCH models describe the conditional variance by relating current volatility to past squared returns and past volatility. This approach is motivated by well-known stylized facts of financial time series: while returns typically display little autocorrelation, squared or absolute returns often exhibit strong and persistent dependence, and volatility tends to cluster over time \citep{mandelbrot1963variation,taylor2011asset}. By explicitly modeling conditional volatility, GARCH-type models provide a natural framework for time-varying financial risk and thus for the forecasting of VaR. However, the increasing availability of high-frequency financial data has created new settings in which risk must be assessed at an intraday level, thereby calling for frameworks that can make fuller use of such data.\\
The increasing availability of high-frequency financial data, together with the growth of high-frequency trading and the demand for timely risk assessment, has intensified the need for methods that capture intraday volatility dynamics. When intraday observations are viewed as realizations of curves over a trading day, Functional Data Analysis (FDA) offers a natural framework for their analysis. FDA studies data that are represented as functions or curves; see, for example, \citet{Ramsay2005}, \citet{hsing2015theoretical}, and \citet{kokoszka2017introduction}.\\
Against this background, recent work has combined ideas from volatility modelling and functional time series analysis. \citet{hormann2013functional} extend the ARCH framework to the functional setting, and \citet{Aue2017} subsequently develop a functional version of the GARCH model. This framework is particularly relevant for high-frequency financial applications, where intraday return dynamics can be represented as curves. In a related application, \citet{RICE20201023} apply the Functional GARCH (FGARCH) model to the forecasting of intraday VaR curves.\\
Despite these developments, the available results are technically demanding and often presuppose substantial background knowledge in functional data analysis and financial econometrics. As a result, the model is not easily accessible to readers seeking a structured introduction to its theoretical background, estimation, and practical use. The present thesis therefore aims to summarize and contextualize the existing literature while also examining aspects that are particularly relevant for applications. In particular, it discusses the relationship between the FGARCH model and closely related alternatives such as the PC-GARCH model, studies finite-sample behavior under heavy-tailed data-generating processes in a Monte Carlo setting, and highlights practical implementation issues. In this sense, the thesis is intended both as an introduction to the FGARCH framework and as an investigation of its use for intraday VaR forecasting.\\
At the same time, the scope of the thesis is deliberately limited. It does not aim to develop new theoretical results for the FGARCH model, nor does it attempt to provide a comprehensive comparison with all competing approaches to multivariate or functional volatility modelling. Instead, the focus lies on presenting and contextualizing the existing theory, comparing the FGARCH and PC-GARCH frameworks, and studying their finite-sample behavior in a Monte Carlo setting with a view toward intraday VaR forecasting. Likewise, although the limitations of VaR are discussed, the thesis does not seek to provide a general assessment of alternative risk measures.\\
Against this background, the aim of this thesis is to examine the theoretical properties and finite-sample performance of the FGARCH model, with a particular focus on forecasting intraday VaR. In particular, this thesis contributes to the existing literature by providing an overview of the theoretical results available for FGARCH models, highlighting selected open problems, and studying their use for intraday VaR forecasting in a Monte Carlo setting. Furthermore, the FGARCH model is related to the principal component GARCH (PC-GARCH) model, since both approaches rely on principal components to reduce dimensionality. Their relationship is examined from a theoretical perspective, and their performance is compared by means of Monte Carlo simulations. Finally, practical issues that arise when applying FGARCH models are discussed, and possible ways to address them are outlined.\\
These objectives can be expressed in terms of the following research questions: What are the main theoretical properties of the FGARCH model, and which difficulties arise when moving from multivariate to functional data? Which estimation strategies for FGARCH models exist, and what are their respective advantages and disadvantages? What practical challenges arise when applying FGARCH models to forecast intraday VaR? How does the FGARCH model perform in finite samples, especially under heavy-tailed data-generating processes? How does its performance compare with that of the PC-GARCH model? What possible extensions of the FGARCH model remain open for future research?\\
The results of this thesis can be summarized as follows:...\\
This thesis proceeds as follows. First, \Cref{subsec:ScalarGARCH} provides a selective overview of the main models and results for univariate and multivariate GARCH processes. Second, \Cref{subsec:FGARCHModel} presents the FGARCH model and discusses its theoretical properties. Third, \Cref{subsec:FGARCHEstimation} introduces two estimation methods for the FGARCH model and examines their consistency and asymptotic distributions. Fourth, \Cref{subsec:ValueAtRisk} presents a mathematical definition of VaR, discusses its main theoretical properties, and reviews important criticisms of the measure. Fifth, \Cref{sec:Simulation} examines the finite-sample performance of the FGARCH model and its implications for forecasting intraday VaR using Monte Carlo simulations. In addition, the FGARCH model is compared with the PC-GARCH model introduced in the preceding sections.\\
Finally, \Cref{sec:Discussion} concludes and discusses extensions of the FGARCH model as well as directions for future research.
\section{Functional GARCH}
\label{sec:FGARCH}
\subsection{(Scalar) GARCH and Multivariate GARCH}
\label{subsec:ScalarGARCH}
Before turning to the FGARCH model, I first consider the scalar version of the GARCH model in both the univariate and multivariate setting.\\
An univariate GARCH$(p,q)$ model can be described by its first two moments:
\begin{definition}[GARCH(p,q) Process]
    A process $(y_t)$ is called a GARCH(p,q) process if its first two conditional moments exist and satisfy:
    \begin{enumerate}
        \item $\mathrm{E}(y_t \,|\, y_u,u<t)=0, \quad \forall t\in\mathbb{Z}$.
        \item There exist constants $\delta,\alpha_i,i=1,...,q$ and $\beta_j,j=1,...,p$ such that\\
        \begin{align}
            \sigma_t^2=\mathrm{Var}(y_t\,|\,y_u,u<t)=\delta+\sum_{i=1}^q\alpha_iy_{t-i}^2+\sum_{j=1}^p\beta_j\sigma_{t-j}^2
        \end{align}
    \end{enumerate}
\end{definition}
To relate the GARCH$(p,q)$ model to more familiar linear time series structures, define the innovation term $\nu_t=y_t^2-\sigma_t^2$ and substitute $\sigma_t^2=y_t^2-\nu_t$ into the volatility equation
\begin{align}
\label{ARMAstructure}
    y_t^2=\delta+\sum_{i=0}^r\overbrace{(\alpha_i+\beta_i)}^{\text{$\kappa_i$}}y_{t-i}^2+\nu_t-\sum_{j=1}^p\beta_j\nu_{t-j}=\delta+\sum_{i=0}^r\kappa_iy_{t-i}^2+\nu_t-\sum_{j=1}^p\beta_j\nu_{t-j}
\end{align}
where $r=\mathrm{max}(p,q)$ with $\alpha_i=0$ ($\beta_j=0$) if $i>q$ ($j>p$). \labelcref{ARMAstructure} shows that the squared process $(y_t^2)$ admits an ARMA$(r,p)$-type representation. This also helps explain why the GARCH model is viewed as a generalization of the ARCH model: just as an ARMA process extends an AR process, the GARCH model extends the ARCH model. Moreover, similarly to the relationship between AR and ARMA models, a GARCH$(p,q)$ process can be represented as an ARCH$(\infty)$ process \citep{francq2019garch}. It should be noted, however, that $(\nu_t)$ is generally not i.i.d. Consequently, the representation in \labelcref{ARMAstructure} cannot be treated as a standard ARMA model when examining the existence of stationary solutions to the GARCH$(p,q)$ process.\\
A more restrictive, yet often useful, version of the GARCH model is the strong GARCH$(p,q)$ process. A strong univariate GARCH$(p,q)$ process is given by:
\begin{gather}
    y_t = \sigma_t\varepsilon_t \\
    \sigma_t^2 = \delta + \sum_{i=1}^q \alpha_i y_{t-i}^2 + \sum_{j=1}^p \beta_j \sigma_{t-j}^2
    \label{Eq2StrongGARCH}
\end{gather}
where $(\varepsilon_t)$ is $\mathrm{i.i.d.}$ In addition, it is often assumed that $\mathbb{E}[\varepsilon_t|\varepsilon_j, j<t]=0$ and $\mathrm{Var}[\varepsilon_t]=1$.\\
In the following, the focus will be on the strong GARCH$(1,1)$ process, since the corresponding functional version is studied in \citet{Aue2017} and applied in \citet{RICE20201023}.
The strong GARCH$(1,1)$ process is the simplest specification within the class of strong GARCH models. Despite its simplicity, it is also one of the most widely used specifications in practice \citep{Zivot2009}.\\
To study the existence of a stationary solution to the strong GARCH$(1,1)$ process, it is useful to rewrite the volatility equation as
\begin{align}
\label{RandomCoeffScalarGARCH}
    \sigma_t^2 = \delta + \gamma_{t-1} \sigma_{t-1}^2, \quad \text{where }\gamma_t=\alpha\varepsilon_t^2+\beta.
\end{align}
This representation shows that the volatility equation of the strong GARCH$(1,1)$ process can be viewed as an autoregressive model with random coefficients. This representation is useful for studying the existence of a strictly stationary and ergodic solution.\\
The following condition guarantees the existence of a strictly stationary, non-anticipative, and ergodic solution to the strong GARCH$(1,1)$ process:
\begin{theorem}[Strict stationarity of the strong GARCH$(1,1)$ process]
\label{stationarityscalarGARCH11}
    If
    \begin{gather}
        -\infty\leq\mathbb{E}[ \log\{\overbrace{\alpha\varepsilon_t^2+\beta}^{\gamma_t}\}]<0,
    \end{gather}
    then the infinite sum
    \begin{gather}
        h_t = \delta \Bigl(1 + \sum_{i=1}^{\infty}\Gamma_{t,i}\Bigr), \quad \text{where }\Gamma_{t,i}=\gamma_{t-1}\cdots\gamma_{t-i}
    \end{gather}
    converges almost surely, and the process $(y_t)$ defined by $y_t=\sqrt{h_t}\varepsilon_t$ is the unique strictly stationary solution of the strong GARCH$(1,1)$ model. This solution is non-anticipative and ergodic.
\end{theorem}
To prove \Cref{stationarityscalarGARCH11}, note that \labelcref{Eq2StrongGARCH} can be solved recursively to obtain $h_t$. Showing that $h_t$ converges involves an application of the Cauchy root test as well as the strong law of large numbers. Uniqueness is then established by contradiction: assuming the existence of another solution, one shows that it must coincide almost surely with the original one.\\
Ergodicity describes the property of a stochastic process that time averages along a single, sufficiently long realization converge to the same values as ensemble averages taken over many realizations. A particularly strong sufficient condition for ergodicity is geometric ergodicity, which requires not only convergence to a unique stationary distribution, but convergence at an exponential rate. It can be shown that $(\sigma_t;t\in \mathbb{Z})$ is geometrically ergodic under suitable regularity conditions. Intuitively, this means that the effect of initial conditions vanishes exponentially fast and that sample averages converge to their population counterparts. Moreover, $(y_t;t\in\mathbb{Z})$ is geometrically $\beta$-mixing, a strong dependence property that implies ergodicity. To prove that the GARCH$(1,1)$ process is geometrically ergodic first note that the volatility equation of the GARCH process can be rewritten into a Markov Chain representation. One can then make use of a Theorem from \citet{FeignTweedie1985}. They show that a Markov chain that is also an aperiodic, irreducible Feller chain is geometrically ergodic if it satisfies a drift condition. Proving that $(y_t;t\in\mathbb{Z})$ is geometrically $\beta$-mixing follows from the geometric ergodicity of $(\sigma_t;t\in\mathbb{Z})$ and the fact that $(\varepsilon_t;t\in\mathbb{Z})$ is i.i.d. $(y_t;t\in\mathbb{Z})$ is not geometrically ergodic since it is not Markovian. The complete proof can be found in \citet{francq2019garch}. A precise mathematical definition of ergodicity is given in the appendix.
%\begin{remark}{Physical intuition behind mixing and ergodicity}
    %A useful intuition for both statistical mixing and ergodicity is obtained from the physical act of stirring different liquids in a glass. For this example, think of the presence of a specific liquid in tiny predefined areas of the glass as a random event. Each observation would be taken after one stir. Each stir would then correspond to iterating the process’s transition mechanism. (Weak) Ergodicity in this context describes the property that each liquid would visit each area of the glass in relation to the dose of the liquid that is contained in the glass and the size of the area after a long time. Geometric ergodicity means that the liquid is observed in each of these areas proportionally to the size of the area and the dose of the liquid exponentially fast.\\
    %Such a process is called mixing when the influence of the past on the future vanishes as the number of stirring steps increases. Different mixing coefficients quantify “how independent” past and future become. That means that for a process that is mixing it is impossible to tell how the liquids were filled into the glass to the start process after sufficient enough stirs. From this example, it can be inferred that geometric ergodicity implies different mixing concepts since after a modest number of stirs the composition is already very close to its long-run law, and any memory of the initial arrangement is forgotten at an exponential rate.\\
    %A physical intuition of both mixing and ergodicity is thus that for such a process where the glass is filled with liquids that are more fluid is likely mixing or geometric ergodic and for such a process where the glass is filled with liquids that are more viscous the process is likely (only) ergodic.
%\end{remark}
\begin{remark}{Physical intuition behind mixing and ergodicity}{rem:IntuitionMixingErgodicity}
    A useful intuition for statistical mixing and ergodicity can be obtained from the physical act of stirring different liquids in a glass. Imagine that the glass is divided into many tiny regions and that we observe which liquid occupies each region after every stir. If we record the configuration after each stir, the resulting sequence of observations can be interpreted as a realization of a stochastic process, while each stir corresponds to one iteration of its transition mechanism.

    Ergodicity intuitively means that, after many stirring steps, the liquids become distributed throughout the glass according to their overall proportions. In other words, the long-run fraction of time that a given liquid is observed in a particular region corresponds to the amount of that liquid in the glass and the size of the region. Geometric ergodicity strengthens this idea by stating that the distribution of liquids across the glass approaches this long-run configuration at an exponential rate.

    The notion of mixing describes how quickly the influence of the past on the future disappears. Mixing coefficients such as the $\alpha$- and $\beta$-mixing coefficients quantify how close observations that are far apart in time are to being statistically independent. In the stirring analogy, after sufficiently many stirs the spatial arrangement of liquids becomes nearly independent of how the liquids were initially poured into the glass. Roughly speaking, the $\alpha$-mixing coefficient measures the remaining dependence between past and future events, while the $\beta$-mixing coefficient provides a stronger measure of how far their joint behavior is from statistical independence. Both therefore quantify how effectively the system "forgets" its earlier states.

    This intuition also illustrates the relationship between ergodicity and mixing. If the liquids mix very efficiently—like highly fluid liquids—then the distribution in the glass quickly approaches its long-run configuration and the dependence between distant observations vanishes rapidly. This corresponds to geometric ergodicity and typically implies strong mixing properties. By contrast, if the liquids are more viscous, the system may still eventually explore all regions of the glass (ergodicity), but the approach to the long-run distribution and the loss of dependence can occur much more slowly.
\end{remark}
%Since in practice, only $y_t$ is observed, the parameters $\delta$,$\alpha$, and $\beta$ have to be estimated. GARCH models are typically estimated using (quasi-)maximum likelihood estimation. (Quasi-)maximum likelihood estimation is preferred for GARCH models over least-squares estimation due to least-squares requiring higher moments and being less efficient than (quasi-)maximum likelihood estimation \citep{francq2019garch}.\\
%The likelihood equations for the GARCH$(1,1)$ process are as follows
%\begin{gather}
   % L_n(\theta;y_1,..,y_n)=\prod_{t=1}^n\frac{1}{\sqrt{2\pi \tilde{\sigma}_t^2}}\exp\left(-\frac{y_t^2}{2\tilde{\sigma}_t^2}\right); \quad \tilde{\sigma}_t^2=\tilde{\sigma}_t^2(\theta)=\delta+\alpha y_{t-1}^2+\beta\tilde{\sigma}_{t-1}^2.
%\end{gather}
%Here $\theta$ is a vector containing the estimates $\delta$, $\alpha$ and $\beta$ and $\tilde{\sigma}_t^2$ is calculated recursively. The true parameters are $\theta_0=(\delta_0,\alpha_0,\beta_0)^\top$.\\
%Taking the logarithm, the likelihood optimization problem can be simplified to minimizing the following equation
%\begin{gather}
%    \mathcal{I}_n=\frac{1}{n}\sum_{t=1}^n \frac{y_t^2}{\tilde{\sigma}_t^2}+\log \tilde{\sigma}_t^2.
%\end{gather}
%Taking the first derivative leads to the following moment condition
%\begin{gather}
   %\frac{1}{n} \sum_{t=1}^n \{y_t^2-\tilde{\sigma}_t^2\}\frac{1}{\tilde{\sigma}_t^4}\frac{\partial \tilde{\sigma}_t^2}{\partial \theta}=0.
%\end{gather}
%From here, parameters can be found by using an optimization algorithm, e.g. the Newton-Raphson method.\\
%It can be shown that the quasi-maximum likelihood estimator is consistent and asymptotically normal with the asymptotic variance of the estimator being equal to the inverse of the Fisher Information matrix when $(\varepsilon_t; t\in\mathbb{Z})$ is Gaussian and the inverse of the adjusted Fisher Information matrix when it is not. The details of the proof are omitted here for the sake of conciseness.\\
Next, I will provide a short exposition of multivariate GARCH processes, as they are very closely related to FGARCH processes in practice. Defining a multivariate GARCH process is not as straightforward as in the univariate case. This is because moving from the univariate to the multivariate setting requires not only the modeling of additional conditional variances, but also the conditional covariances between the components. Without further restrictions, the resulting number of parameters quickly becomes too large for reliable estimation and asymptotic analysis. Therefore, the conditional covariance matrix $H_t$ follows an explicit specification. In the following, I will present two of these specifications since they are closely connected to the FGARCH model. Throughout this discussion, I will assume that $y_t,\varepsilon_t\in \mathbb{R}^m$ and that $\varepsilon_t$ is an i.i.d. vector.\\
First, I consider the PC-GARCH model, which is defined by
\begin{gather}
\label{PCGARCHyequation}
    y_t = H_t^{\frac{1}{2}}\varepsilon_t,\\
    \label{PCGARCHHequation}
    H_t = P \Lambda_t P^{\top}.
\end{gather}
Here, $P$ is an $m\times m$ orthogonal matrix, while $\Lambda_t = \mathrm{diag}(\lambda_{1,t},\lambda_{2,t},\ldots,\lambda_{m,t})$ is an $m\times m$ diagonal matrix. The matrix $P$ is typically estimated using Principal Component Analysis, and the diagonal elements of $\Lambda_t$ are assumed to follow univariate GARCH$(1,1)$ processes:
\begin{gather}
\label{PCGARCHlambdaequation}
    \lambda_{i,t} = \delta_i + \alpha_i f_{i,t-1}^{2} + \beta_i \lambda_{i,t-1}, \quad i = 1, \ldots, m.
\end{gather}
In this specification, $f_{i,t}$ denotes the $i$-th component of the transformed series $f_t = P^\top y_t$. The parameters in \Cref{PCGARCHlambdaequation} are typically estimated by quasi-maximum likelihood. Alternative specifications for $\lambda_{i,t}$ are also possible.\\
From the representation in \labelcref{PCGARCHHequation}, it follows directly that the PC-GARCH model decomposes the conditional covariance matrix $H_t$ into a set of constant eigenvectors and time-varying eigenvalues. As will be discussed in \Cref{sec:Simulation}, the PC-GARCH model is closely related to a discretized version of the FGARCH model, in which the underlying functions are observed on a finite grid.\\
The second multivariate GARCH specification considered is the Constant Conditional Correlation (CCC) GARCH model, which is defined as
\begin{gather}
    y_t = H_t^{\frac{1}{2}}\varepsilon_t,\\
    \label{CCCGARCHHequation}
    H_t = D_t R D_t^{\top}.
\end{gather}
Here, $D_t=\text{diag}(\sqrt{h_{1,t}}, \sqrt{h_{2,t}}, \ldots, \sqrt{h_{m,t}})$ is an $m \times m$ diagonal matrix and $R$ is an $m \times m$ constant correlation matrix. The vector of conditional variances $h_t$ is assumed to follow a GARCH(1,1)-type specification given by
\begin{gather}
\label{CCCGARCHsmallhequation}
    h_t = \delta + A (y_{t-1} \odot y_{t-1}) + B h_{t-1},
\end{gather}
where $\odot$ denotes the elementwise (Hadamard) product, $\delta$ is an $m$-dimensional vector, and $A$ and $B$ are $m \times m$ parameter matrices. \labelcref{CCCGARCHHequation} shows that $H_t$ consists of time-varying, variable-specific conditional variances combined with a constant correlation structure. In addition, \Cref{CCCGARCHsmallhequation} shows that the variable-specific conditional variances evolve dynamically, depending not only on each variable’s own past but also on the past of the other variables. As will be discussed in \Cref{subsec:FGARCHEstimation}, once the functional parameters are projected onto a finite set of basis functions, estimation of the FGARCH model reduces to that of a semi-strong CCC-GARCH model with $R = \mathbb{I}$, where $\mathbb{I}$ denotes the identity matrix.\\
Since in practice only the process $(y_t)_{t\in\mathbb{Z}}$ is observed, the parameters of any univariate or multivariate GARCH model have to be estimated. GARCH models are typically estimated using (quasi-)maximum likelihood estimation. For GARCH models, (quasi-)maximum likelihood estimation is generally preferred to least-squares estimation, since least-squares methods require stronger moment conditions and are typically less efficient \citep{francq2019garch}. In the following, I will illustrate the estimation procedure for the univariate GARCH$(1,1)$ model. Analogous arguments apply to other GARCH specifications.\\
The Gaussian quasi-likelihood for the GARCH$(1,1)$ model is given by
\begin{gather}
    L_n(\underline{\theta};y_1,..,y_n)=\prod_{t=1}^n\frac{1}{\sqrt{2\pi \tilde{\sigma}_t^2}}\exp\left(-\frac{y_t^2}{2\tilde{\sigma}_t^2}\right); \quad \tilde{\sigma}_t^2=\tilde{\sigma}_t^2(\underline{\theta})=\underline{\delta}+\underline{\alpha} y_{t-1}^2+\underline{\beta}\tilde{\sigma}_{t-1}^2.
\end{gather}
Here, $\underline{\theta}=(\underline{\delta},\underline{\alpha},\underline{\beta})^\top$ denotes the parameter vector, and $\tilde{\sigma}_t^2$ is computed recursively. The true parameter vector is denoted by $\theta=(\delta,\alpha,\beta)^\top$.\\
Taking logarithms and omitting constants, the optimization problem is equivalent to minimizing
\begin{gather}
    \mathcal{I}_n=\frac{1}{n}\sum_{t=1}^n\left( \frac{y_t^2}{\tilde{\sigma}_t^2}+\log \tilde{\sigma}_t^2\right).
\end{gather}
The parameter estimator is then obtained numerically by minimizing this criterion, for example by using a Newton-Raphson algorithm.\\
Under suitable regularity conditions, the quasi-maximum likelihood estimator is consistent and asymptotically normal. If the innovations are Gaussian, its asymptotic variance is given by the inverse Fisher information matrix. If the innovations are non-Gaussian, the asymptotic variance instead involves the inverse of an adjusted information matrix. The details of the proof are omitted here for the sake of conciseness.
\subsection{Model}
\label{subsec:FGARCHModel}
This section introduces the FGARCH model and discusses its key theoretical properties. I begin with the definition of the model.\\
A FGARCH process is defined as follows:
\begin{definition}[FGARCH(p,q) Process]
\label{fGARCHpq}
A sequence of random functions $(y_i(t))_{i\in \mathbb{Z}}$ is called a FGARCH process of order $(p,q)$ if it satisfies the following equations
    \begin{gather}
        y_i(t) = \sigma_i(t)\varepsilon_i(t)\\
        \sigma_i^2(t) = \delta(t)+\sum_{j=1}^q\alpha_jy_{i-j}^2(t)+\sum_{k=1}^p\beta_k\sigma_{i-k}^2(t).
    \end{gather}
\end{definition}
Here, $\delta(t)\in L^2[0,1]$ is a non-negative function, and $(\alpha_j)_{j\in (1,...,p)}$ and $(\beta_k)_{k \in (1,...,q)}$ are bounded integral operators in $L^2[0,1]^2$ given by $(\alpha_jx)(t)=\int \alpha_j(t,s)x(s)ds$ and $(\beta_kx)(t)=\int \beta_k(t,s)x(s)ds$ that map non-negative functions to non-negative functions for $x\in L^2[0,1]$. By convention, I restrict the intraday time variable to $t\in[0,1]$. This is without loss of generality since data that do not follow this convention can be converted into this format. Accordingly, $\int_0^1$ will be abbreviated by $\int$. The second time variable $i$ denotes the trading day. The innovations $(\varepsilon_i(t))_{i\in\mathbb{Z}}$ are i.i.d. random functions.\\
From \Cref{fGARCHpq}, it is clear that the FGARCH$(p,q)$ process is the functional analogue of the strong (scalar) GARCH process. As a consequence, the model admits an ARMA-type representation, just as in the scalar case discussed in \Cref{ARMAstructure}.\\
Following \citet{Aue2017} and \citet{RICE20201023}, I focus on the FGARCH$(1,1)$ specification, which is the functional counterpart of the benchmark strong (scalar) GARCH$(1,1)$ model. As discussed previously, despite its simplicity the strong GARCH$(1,1)$ process is also one of the most widely used specifications in practice. For this reason, the FGARCH$(1,1)$ model is a natural starting point in the functional setting. In this case, the FGARCH$(p,q)$ equations reduce to
\begin{gather}
    \label{FunctionalGARCH11}
    y_i(t) = \sigma_i(t)\varepsilon_i(t)\\
    \label{FunctionalGARCH11volatility}
    \sigma_i^2(t) = \delta(t)+\alpha y_{i-1}^2(t)+\beta\sigma_{i-1}^2(t).
\end{gather}
In the following, conditions for the existence of a stationary solution to the FGARCH$(1,1)$ process are discussed. \citet{Aue2017} derives these results both for the space $L^2[0,1]$ as well as for the space of continuous functions $C[0,1]$. I restrict attention to the $L^2[0,1]$ setting, since the corresponding results for $C[0,1]$ are of essentially the same form. The main difference is that the innovation functions and the intercept function $\delta(t)$ are assumed to belong to $C[0,1]$, whereas the kernel of the integral operators remain in $L^2[0,1]^2$, so that the resulting operators map $C[0,1]$ into $C[0,1]$.\\
Using notation analogous to the scalar case, the volatility equation can be written as
\begin{gather}
\label{FGARCH11volatilityalternative}
    \sigma_i^2(t) = \delta(t) + \gamma_{i-1}\sigma_{i-1}^2(t),
\end{gather}
where $\gamma_i$ is the random integral operator with kernel
\begin{gather}
    \gamma_i(t,s)=\alpha(t,s)\varepsilon_i^2(s)+\beta(t,s).
\end{gather}
The theorem below assumes that the kernels of $\alpha$ and $\beta$ belong to $L^2([0,1]^2)$ and that $\varepsilon_0\in L^4[0,1]$ almost surely. These assumptions imply that $\gamma_i\in L^2([0,1]^2)$ with probability $1$. Consequently, $\gamma_i$ defines a bounded random integral operator on $L^2[0,1]$; see \citet{riesz1990functional}. Writing the volatility recursion in this form is convenient, since the existence of a strictly stationary solution can then be characterized through properties of the random operator $\gamma_i$ and, in particular, its Hilbert--Schmidt norm. The following theorem provides sufficient conditions for the existence of a strictly stationary solution.
\begin{theorem}[Strict stationarity of the FGARCH$(1,1)$ process]
\label{StationarityFGARCH11}
    Let $\delta(t)\in L^2[0,1]$ be a non-negative function, and let $\alpha$ and $\beta$ be integral operators with kernels in $L^2[0,1]^2$ that map non-negative functions to non-negative functions. Let $\{\varepsilon_i(t):i\in \mathbb{Z}\}$ be i.i.d. random functions such that $\varepsilon_0\in L^{4}[0,1]$ almost surely.
    \begin{enumerate}[label=(\roman*)]
        \item
        \label{FGARCHStationarySolution}
        If
            \begin{gather}
            \label{FGARCH11StationaryCondition}
                 -\infty \leq \mathbb{E}\left[\log \left\| \gamma_0 \right\|_\mathcal{S}\right] < 0
            \end{gather}
            then the infinite sum
            \begin{gather}
            \label{FGARCH11StationarySolution}
                h_i(t)=\delta(t)+\sum_{k=1}^\infty \Gamma_{i,k}\delta(t)
            \end{gather}
            converges almost surely, and the process $(y_i(t))$, defined by $y_i(t)=\sqrt{h_i(t)}\varepsilon_i(t)$, is the unique strictly stationary solution to the FGARCH$(1,1)$ equations in \labelcref{FunctionalGARCH11} and \labelcref{FunctionalGARCH11volatility}. This solution is non-anticipative and ergodic.
        \item
        \label{FGARCHMoments}
        If there is $\nu>0$ such that 
        \begin{gather}
        \label{FGARCH11StationarityMomentsCondition}
            \mathbb{E}\left[\left\| \gamma_0 \right\|_\mathcal{S}^\nu\right] < 1
        \end{gather}
        then there exists a unique, strictly stationary, and non-anticipative solution in $L^2[0,1]$ to the FGARCH$(1,1)$ model and $\mathbb{E}\left[\left\| \sigma_0^2 \right\|_2^\nu\right] < \infty$.
    \end{enumerate}
\end{theorem}
Here $\left\|\right\|_\mathcal{S}$ is the Hilbert-Schmidt norm defined as
\begin{gather}
    \left\| \gamma_0 \right\|_\mathcal{S}=\left(\int\!\!\int \gamma_0^2(t,s)\,dt\,ds\right)^{1/2}
\end{gather}
and $\Gamma_{i,k}$ denotes the random integral operator
\begin{gather}
    (\Gamma_{i,k}x)(t)=\int \cdots \int\gamma_{i-1}(t,s_1)\gamma_{i-2}(s_1,s_2)\cdots\gamma_{i-k}(s_{k-1},s_k)x(s_k)\,ds_k\cdots ds_2\,ds_1.
\end{gather}
Due to the close resemblance between \labelcref{FGARCHStationarySolution} and \Cref{stationarityscalarGARCH11}, the proof of \labelcref{FGARCHStationarySolution} follows a very similar line of argument. First, $h_i(t)$ is obtained by recursively iterating \labelcref{FGARCH11volatilityalternative}. Second, almost sure convergence of $h_i(t)$ is established by combining the Cauchy--Schwarz inequality with the strong law of large numbers applied to the logarithms of the operator norms. Finally, the uniqueness of the solution is shown by exploiting the recursive structure of \labelcref{FGARCH11volatilityalternative}. Assuming the existence of an alternative solution implies that both solutions converge to the same limit and must therefore coincide, yielding a contradiction. The proof of \labelcref{FGARCHMoments} proceeds by first establishing inequalities between $\|\sigma_0^2\|_2$ and the series $\sum_{k=0}^\infty \|\Gamma_{i,k}\delta\|_2$ using the Cauchy--Schwarz inequality. Applying Minkowski's inequality then yields the claimed result on the existence of moments in \labelcref{FGARCHMoments}.\\
The full proof of \Cref{StationarityFGARCH11} can be found in the Appendix.\\
%My previous version:
%Due to the strong similarity of this theorem to \ref{stationarityscalarGARCH(1,1)}, proving \ref{StationarityFGARCH(1,1)} follows a similar idea. Firstly, $h_i(t)$ is derived by recursively solving (\ref{FGARCH(1,1)volatilityalternative}). Secondly, to prove the convergence of $h_i(t)$ an application of the Cauchy-Schwarz inequality as well as the strong law of large numbers is needed. Lastly, to prove the uniqueness of the solution, the recursive structure of (\ref{FGARCH(1,1)volatilityalternative}) is used. Assuming that another solution exists results in the two solutions converging to the same limit and therefore both solutions coincide resulting in a contradiction.
%Both \citet{Cerovecki2019} and \citet{Kuehnert2020} extend the stationarity result in (\ref{FGARCH(1,1)StationaryCondition}). \citet{Cerovecki2019} provide sufficient conditions under which the FGARCH model is stationary for the $L^2$ space for any length $p,q$. These conditions are sufficient but not necessary, as stating necessary conditions is more complicated in the functional setting since norms are not equivalent in the infinite dimensional setting. To do so, they make use of the top Lyapunov exponent \citep{Kingman1973}. They show that when applying their result to the FGARCH$(1,1)$ model a milder condition than (\ref{FGARCH(1,1)StationaryCondition}) can be established. However, to do so, they require a different assumption on $\varepsilon_0$. \citet{Kuehnert2020} further expands conditions for stationarity for the FGARCH model to $L^p$ where $p\in[1,\infty)$ and Banach spaces.\\ My version
Both \citet{Cerovecki2019} and \citet{Kuehnert2020} extend the stationarity result in \labelcref{FGARCHStationarySolution}. \citet{Cerovecki2019} derive sufficient conditions for the existence of a unique strictly stationary and non-anticipative FGARCH$(p,q)$ solution in $L^2$ for arbitrary orders $p$ and $q$. Their approach is based on the top Lyapunov exponent \citep{Kingman1973}. They also note that, in contrast to the finite-dimensional case, establishing necessity is substantially more difficult in the functional setting because norms are not equivalent in infinite-dimensional spaces. When specialised to the FGARCH$(1,1)$ specification, their general result yields a milder sufficient condition than \labelcref{FGARCH11StationaryCondition}, although this comes at the cost of imposing a different assumption on the innovation process.\citet{Kuehnert2020} further generalizes both the stationarity results and the result on the existence of moments in \labelcref{FGARCHMoments} to $L^p$ spaces with $p \in [1,\infty)$, and more generally to separable Banach spaces of functions on $[0,1]$ endowed with the supremum norm and closed under pointwise multiplication.\\
To establish the ergodicity of the stationary solution, the following corollary is useful.
\begin{corollary}
The stationary non-anticipative solution established in \Cref{StationarityFGARCH11} yields the following corollary:
    \begin{enumerate}[label=(\roman*)]
        \item
        \label{FirstPointCorollary}
        There is a functional $g$ such that
        \begin{gather}
            \sigma_i^2=g(\varepsilon_{i-1},\varepsilon_{i-2},...).
        \end{gather}
        As a consequence, $(\sigma_i^2(t);i\in\mathbb{Z})$ and $(y_i(t);i\in \mathbb{Z})$ are Bernoulli shifts. This follows directly from \labelcref{FGARCH11StationarySolution}.
        \item
        \label{SecondPointCorollary}
        \labelcref{FirstPointCorollary} implies that both $(\sigma_i^2(t); i\in \mathbb{Z})$ and $(y_i(t);i\in\mathbb{Z})$ are ergodic.
    \end{enumerate}
\end{corollary}
%Ergodicity of Bernoulli shifts is a well established result and follows from \citet{stout1974almost}.\\
%Unlike in the scalar GARCH model, establishing geometric ergodicity and $\beta$-mixing for the FGARCH model is substantially more difficult. This is due to the lack of densities for infinite dimensional data, as well as the need for additional constraints on the functional parameters $\alpha$ and $\beta$ to establish continuity, like in the proof of geometric ergodicity for the scalar GARCH model. Furthermore, showing that the GARCH process is an irreducible Markov chain is also difficult due to the infinite dimensionality of the data. Establishing small sets is equally complicated. However, ergodicity alone is not sufficient for the derivation of important statistical results like central limit theorems. That is because information on how fast dependence decays is needed to establish the convergence of sample estimates to real values in distribution. Since ergodicity is not sufficient for the application of the functional central limit theorem, \citet{Aue2017} derives the following result. My previous version
Ergodicity of Bernoulli shifts is a well established result and follows from \citet{stout1974almost}.\\
In the functional setting, arguments of the type used in the scalar case become much harder to apply. This is, for example, due to the fact that infinite-dimensional observations do not admit densities in the same way as finite-dimensional random variables. In addition, stronger assumptions on the functional parameters $\alpha$ and $\beta$ may be needed to establish continuity properties of the functional recursion, similar to the role played by continuity in the proof of geometric ergodicity for the scalar GARCH model. Furthermore, showing that the FGARCH process forms an irreducible Markov chain is difficult in the infinite-dimensional setting, and establishing small sets is equally complicated. However, ergodicity alone is not sufficient for the derivation of important statistical results like central limit theorems. That is because information on how fast dependence decays is needed to establish convergence in distribution of suitable sample statistics. Since ergodicity is not sufficient for the application of the functional central limit theorem, \citet{Aue2017} derives the following result.
\begin{corollary}
\label{SecondCorollaryModelSection}
    \begin{enumerate}[label=(\Roman*)]
        \item
        \label{FGARCH11CLTCondition}
        For $\ell\in \mathbb{N}$ and $j,i\in\mathbb{Z}$, let $\varepsilon_{i,j}^{(\ell)}(t)$ be independent copies of $\varepsilon_0(t)$ and define
        \begin{gather}
            \sigma_{i,\ell}^2(t) = g\left(\varepsilon_i(t),...,\varepsilon_{i-\ell}(t),\varepsilon_{i-\ell-1,i-\ell}^{(i)}(t),\varepsilon_{i-\ell-2,i-\ell}^{(i)}(t),... \right).
        \end{gather}
        If \labelcref{FGARCH11StationarityMomentsCondition} is satisfied, then there exists a constant $0<\rho<1$ such that
        \begin{gather}
        \label{Lpmequation}
            \mathbb{E}\left[\left\|\sigma_i^2-\sigma_{i,\ell}^2 \right\|_2^\nu\right]=\mathcal{O}(\rho^\ell).
        \end{gather}
        \item From \labelcref{FGARCH11CLTCondition}, it follows that $(\sigma_i^2(t);i\in\mathbb{Z})$ and $(y_i(t);i\in \mathbb{Z})$ are weakly dependent processes.
    \end{enumerate}
\end{corollary}
\labelcref{FGARCH11CLTCondition} expresses geometric $L^\nu$-m-approximability with respect to the $L^2$-norm. The concept of $L^p$-m-approximability for functional data was first introduced by \citet{Lpmapproxibility}. A process that is geometrically $L^p$-m-approximable can be approximated by an m-dependent process at a geometric rate. A stochastic process $(X_n)$ is called m-dependent if observations that are more than $m$ time steps apart are independent. More precisely, for any time index $k$, the collection of random variables $\left\{X_n:n\leq k \right\}$ is independent of the collection $\left\{X_n:n\geq k+m \right\}$. In other words, the past up to time $k$ has no influence on the future starting from time $k+m$. Formally, $X_n$ is m-dependent if, for any integer $k$, the $\sigma$-algebra $\mathcal{F}_k^-=\sigma\left\{...,X_{k-2},X_{k-1},X_k \right\}$ generated by the past observations is independent of the $\sigma$-algebra generated by the future observations $\mathcal{F}_{k+m}^+=\sigma\left\{ X_{k+m},X_{k+m+1},X_{k+m+2},... \right\}$. \citet{Aue2017} use $L^p$-$m$-approximability to show that the empirical eigenfunctions computed from the empirical covariance kernel of $y_i^2(t)$ converge to the corresponding eigenfunctions of the true covariance kernel of $y_i^2(t)$. The use of this result will be discussed in \Cref{subsec:FGARCHEstimation}. They also establish that, if \labelcref{Lpmequation} holds for some $\nu \geq 2$, this weak-dependence condition is strong enough to derive a functional central limit theorem; see \citet{BERKES2013385} and \citet{Jirak2013} for background on the weak invariance principles underlying such results.\\
Finally, \citet{Kuehnert2020} extends the $L^\nu$-m-approximability result in \labelcref{Lpmequation} to FGARCH processes of arbitrary orders $p$ and $q$, as well as to $L^p$ spaces with $p \in [1,\infty)$, and more generally to separable Banach spaces of functions on $[0,1]$ endowed with the supremum norm and closed under pointwise multiplication.
\subsection{Estimation}
\label{subsec:FGARCHEstimation}
%\begin{itemize}
%    \item Projection onto finite dimensional basis
%    \item Least Squares Minimization problem
%    \item Solving Least Squares (Connection to uni-/ multivariate GARCH models)
%    \item Asymptotic results
%    \item (Quasi-Maximum Likelihood Estimation?)
%\end{itemize}
%In this next section, the estimation of a FGARCH$(1,1)$ model will be discussed. To this end, it is assumed that there exists a finite sample of $N$ observed functions $(y_i(t))_{i\in[1,N]}$. In practice, functional objects are typically observed on a fine grid, however, here it is assumed that the functional objects themselves are observed or can be retrieved from the data.\\ My Version before refining with ChatGPT.
In this section, the estimation of an FGARCH$(1,1)$ model will be discussed. I consider a setting in which a finite sample of $N$ functional observations $\{y_i(t)\}_{i=1}^N$ is available. Although functional data are commonly recorded on a dense discrete grid in practical applications, I assume here that the underlying curves are directly observed or can be reliably reconstructed from the data.\\
Because the objects to be estimated are infinite-dimensional, direct estimation from a finite sample is generally ill-posed and therefore requires some form of dimension reduction. For this reason, \citet{Aue2017} impose a standard finite-dimensional approximation on the intercept function $\delta(t)$ and on the integral operators $\alpha$ and $\beta$. In particular, they assume that all three objects can be approximated using a fixed number $M$ of orthonormal basis functions $\Phi_M = \{\varphi_1(t), \ldots, \varphi_M(t)\}$ defined on $[0,1]$. Under this assumption, $\delta(t)$ is represented as a finite linear combination of these basis functions, while the operators $\alpha$ and $\beta$ are represented through their kernel functions using the same basis:
\begin{align}
\label{basisdecompositiondelta}
    \delta(t) &= \sum_{m=1}^M d_m \varphi_m(t), \\
    \alpha(t,s) &= \sum_{m,m'=1}^M a_{m,m'} \varphi_m(t)\varphi_{m'}(s), \\
    \label{basisdecompositionbeta}
    \beta(t,s) &= \sum_{m,m'=1}^M b_{m,m'} \varphi_m(t)\varphi_{m'}(s).
\end{align}
From this representation, it is easy to see that estimating the functional objects $\delta(t)$, $\alpha$ and $\beta$ simplifies to estimating $\{d_1,\ldots,d_M,a_{1,1},\ldots,a_{M,M},b_{1,1},\ldots,b_{M,M}\}$. To further simplify notation, \citet{Aue2017} define the following quantities:
\begin{align}
\label{MatricesdAB}
    \mathbf{d}=\begin{pmatrix}
        d_1\\
        \vdots\\
        d_M
    \end{pmatrix}; \quad
    \mathbf{A}=\begin{pmatrix}
        a_{1,1} & \ldots & a_{1,M}\\
        \vdots & \ddots & \vdots\\
        a_{M,1} & \ldots & a_{M,M}
    \end{pmatrix}; \quad
    \mathbf{B}=\begin{pmatrix}
        b_{1,1} & \ldots & b_{1,M}\\
        \vdots & \ddots & \vdots\\
        b_{M,1} & \ldots & b_{M,M}
    \end{pmatrix}
\end{align}
To estimate $\mathbf{d}$, $\mathbf{A}$, and $\mathbf{B}$, the squared observations $(y_i^2(t))_{i\in[1,N]}$ and conditional volatility functions $(\sigma_i^2(t))_{i\in[1,N]}$ are projected onto $\Phi_M$. This yields the M-dimensional vectors $\mathbf{y}_i^{(2)}=\begin{pmatrix}
    y_{i,1}^{(2)} & \ldots & y_{i,M}^{(2)}
\end{pmatrix}^\top$ and $\mathbf{s}_i^{(2)}=\begin{pmatrix}
    s_{i,1}^{(2)} & \ldots & s_{i,M}^{(2)}
\end{pmatrix}^\top$ whose components are given by the inner products with the basis functions in $\Phi_M$ $y_{i,m}^{(2)}=\langle y_i^2(t),\varphi_m(t)\rangle$, $s_{i,m}^{(2)}=\langle \sigma_i^2(t),\varphi_m(t)\rangle$. Under the finite-dimensional parametrisation introduced above, the volatility recursion can then be written in terms of the projected coefficient vectors as
\begin{align}
\label{FGARCH11MatrixForm}
    \mathbf{s}_i^{(2)}=\mathbf{d}+\mathbf{Ay}_{i-1}^{(2)}+\mathbf{Bs}_{i-1}^{(2)}.
\end{align}
Solving this equation recursively obtains an infinite-order representation for $\mathbf{s}_i^{(2)}$ that only depends on $\mathbf{y}_{i}^{(2)}$
\begin{align}
\label{Matrixrepinftrealparameters}
    \mathbf{s}_i^{(2)}=\sum_{\ell=1}^{\infty} \mathbf{B}^{\ell-1}\mathbf{d}+\sum_{\ell=1}^{\infty}\mathbf{B}^{\ell-1}\mathbf{Ay}_{i-\ell}^{(2)}.
\end{align}
Since the true parameter vector $\mathbf{\theta}=(\mathbf{d}^{\top},\mathbf{A}^\top,\mathbf{B}^\top)^\top$ is unknown, \labelcref{Matrixrepinftrealparameters} is expressed at an arbitrary parameter value $\mathbf{\underline{\theta}}=(\underline{\mathbf{d}}^{\top},\underline{\mathbf{A}}^\top,\underline{\mathbf{B}}^\top)^\top\in\mathbf{\Theta}\subset\mathbb{R}^{M+2M^2}$
\begin{align}
    \mathbf{\tilde{s}}_i^{(2)}=\mathbf{\tilde{s}}_i^{(2)}(\underline{\theta})=\sum_{\ell=1}^{\infty}\underline{\mathbf{B}}^{\ell-1}\underline{\mathbf{d}}+\sum_{\ell=1}^{\infty}\underline{\mathbf{B}}^{\ell-1}\underline{\mathbf{A}}\mathbf{y}_{i-\ell}^{(2)}.
\end{align}
Since in practice only a finite time frame of the series can be observed, $\mathbf{s}_i^{(2)}$ is approximated by $\mathbf{\hat{s}}_i^{(2)}$
\begin{align}
    \mathbf{\hat{s}}_i^{(2)}=\mathbf{\hat{s}}_i^{(2)}(\underline{\theta})=\sum_{\ell=1}^{i-1}\underline{\mathbf{B}}^{\ell-1}\underline{\mathbf{d}}+\sum_{\ell=1}^{i-1}\underline{\mathbf{B}}^{\ell-1}\underline{\mathbf{A}}\mathbf{y}_{i-\ell}^{(2)}.
\end{align}
The objective, thus, has to be to choose an estimator $\mathbf{\underline{\hat{\theta}}}_n$ such that $\mathbf{\hat{s}}_i^{(2)}$ is close to $\mathbf{s}_i^{(2)}$. Since $\mathbf{s}_i^{(2)}$ is unobservable, \citet{Aue2017} impose
\begin{gather}
\label{UnitVarianceepsilonfunction}
    \mathbb{E}[\varepsilon_i^2(t)]=1; \quad \forall t\in[0,1].
\end{gather}
This is without loss of generality since this condition can always be fulfilled by an appropriate way of scaling. Under this condition, $\mathbb{E}[y_i^{2}(t)|\mathcal{F}_{i-1}]=\sigma_i^{2}(t)$ and therefore, after projection onto $\Phi_M$, $\mathbb{E}[y_i^{(2)}|\mathcal{F}_{i-1}]=s_i^{(2)}$ where $\mathcal{F}_{i-1}$ is the sigma-algebra generated by all $\varepsilon_\ell:\ell<i$, $\mathcal{F}_{i-1}=\sigma(\varepsilon_\ell:\ell<i)$. As a consequence, \citet{Aue2017} define the following least squares estimator
\begin{align}
\label{ObjectiveFunctionLSFGARCH}
    \underline{\hat{\theta}}_n=\arg \min\{S_n(\underline{\theta})=\sum_{i=2}^n\{y_i^{(2)}-\hat{s}_i^{(2)}(\underline{\theta}) \}^{\top}\{y_i^{(2)}-\hat{s}_i^{(2)}(\underline{\theta}) \}:\underline{\theta}\in\mathbf{\Theta}\}.
\end{align}
In the following, I discuss the strong consistency and asymptotic distribution of the least-squares estimator $\underline{\hat{\theta}}_n$. The required assumptions are summarized below in a slightly condensed form following \citet{Aue2017}.
\begin{assumption}
\label{Assumption1}
    It is assumed that
    \begin{enumerate}[label=(A\arabic*)]
        \item $\mathbf{y}_1^{(2)}$ is not measurable with respect to $\mathcal{F}_0$,
        \item $\mathbf{\Theta}$ is a compact set and $\theta$ is in the interior of $\mathbf{\Theta}$ and
        \item there are $0<c_1$ and $c_2<1$ such that $c_1\leq \left|\det(\underline{\mathbf{A}})\right|$ and $\left\| \mathbf{B} \right\|<1$ for all $\underline{\theta}=(\underline{\mathbf{d}}^\top,\underline{\mathbf{A}}^\top,\underline{\mathbf{B}}^\top)\in\mathbf{\Theta}$.
        \label{Ainvertibility}
    \end{enumerate}
\end{assumption}
The second part is a standard compactness and interiority condition on the parameter space. The third part imposes uniform identification and stability restrictions over $\Theta$. Consequently, it guarantees that $\underline{\mathbf{A}}$ is non-singular and that $\underline{\mathbf{B}}$ is uniformly contractive for all $\underline{\theta}\in \Theta$. In particular, these properties hold at the true parameter value $\theta$.\\
Using this assumption, the strong consistency of $\underline{\hat{\theta}}_n$ is established in the following Theorem.
\begin{theorem}
\label{StrongConstistencyFGARCH11}
    Assume that $(y_i(t);i\in\mathbb{Z})$ follows a FGARCH$(1,1)$ process as defined in \labelcref{FunctionalGARCH11} and \labelcref{FunctionalGARCH11volatility}, that $\delta(t)$ is a non-negative function, and that $\alpha$ and $\beta$ are integral operators that map non-negative functions to non-negative functions, that \labelcref{FGARCH11StationarityMomentsCondition} holds with $\nu=1$, that $\mathbb{E}\left[\left\|\varepsilon_0\right\|_2^2\right]<\infty$, and that \labelcref{UnitVarianceepsilonfunction} holds. Then under \Cref{Assumption1} $\underline{\hat{\theta}}_n$ is strongly consistent, meaning that
    \begin{gather}
        \underline{\hat{\theta}}_n \rightarrow \theta \quad \mathrm{a.s.}
    \end{gather}
    as $n\rightarrow\infty$.
\end{theorem}
To examine the asymptotic distribution of $\underline{\hat{\theta}}_n$, the following quantities are useful
\begin{gather}
    \mathbf{H_0}=\mathbb{E}\left[\frac{\partial \tilde{\mathbf{s}}_0^{(2)}(\theta)}{\partial\underline{\theta}}\right], \quad \quad \mathbf{J_0}=\mathbb{E}\left[\left\{\mathbf{y}^{(2)}-\mathbf{s}_0^{(2)}\right\}\left\{\mathbf{y}^{(2)}-\mathbf{s}_0^{(2)}\right\}^\top\right],\\
    \mathbf{Q_0}=\mathbb{E}\left[\left(\frac{\partial \tilde{\mathbf{s}}_0^{(2)}(\theta)}{\partial\underline{\theta}}\right)^\top\left(\frac{\partial \tilde{\mathbf{s}}_0^{(2)}(\theta)}{\partial\underline{\theta}}\right)\right].
\end{gather}
Under the conditions stated in \Cref{StrongConstistencyFGARCH11}, the matrices $H_0$ and $Q_0$ are well defined. Moreover, if $\mathbb{E}[\left\|\sigma_0^2\right\|_2^4]<\infty$ and $\mathbb{E}[\left\|\varepsilon_0^2\right\|_2^4]<\infty$, the matrix $J_0$ exists. Here, $\mathbf H_0$ is the expected first-derivative matrix of the fitted volatility coefficients with respect to the parameter vector, while $\mathbf J_0$ is the covariance matrix of the residual term $\mathbf y_0^{(2)}-\mathbf s_0^{(2)}$. Thus, $\mathbf H_0$ captures the local sensitivity of the model with respect to the parameters, whereas $\mathbf J_0$ measures the variability of the part of the projected squared observations that is not explained by the conditional volatility. The matrix $\mathbf Q_0$ is the second-moment matrix of the derivative of $\tilde{\mathbf s}_0^{(2)}(\theta)$ with respect to the parameter vector. In this sense, it plays the role of a curvature matrix for the least-squares criterion. More specifically, as shown in the Appendix, the second derivative of the least-squares objective converges to $2\mathbf Q_0$ in the proof of the asymptotic result. Following the standard theory of $M$-estimators, it is therefore additionally assumed that $\mathbf Q_0$ is non-singular. This guarantees invertibility and ensures that the objective function is locally well behaved around the true parameter.
\begin{theorem}
\label{AsymptoticDistroFGARCH11estimator}
Assume that the conditions of \Cref{StrongConstistencyFGARCH11} are satisfied, that $\mathbf{Q_0}$ is non-singular, and that $\mathbb{E}\!\left[\|y_0^2\|_2^4\right]<\infty$. Then, as $n\to\infty$,
\begin{gather}
\sqrt{n}\,(\underline{\hat{\theta}}_n-\theta)
\;\overset{\mathcal{D}}{\longrightarrow}\;
\mathcal{N}\!\left(0,\,
\mathbf{Q_0}^{-1}\mathbf{H_0}^\top\mathbf{J_0H_0Q_0}^{-1}\right),
\end{gather}
where the limiting distribution is a $(2M^2+M)$-dimensional normal distribution.
\end{theorem}
%The proof of \ref{AsymptoticDistroFGARCH(1,1)estimator} uses the (componentwise) mean value theorem. First, define $\hat{Z}_n(\underline{\theta})=\frac{1}{n}\sum_{i=2}^n\left\{ y_i^{(2)}-\hat{s}_i^{(2)}(\underline{\theta})\right\}^\top \left\{ y_i^{(2)}-\hat{s}_i^{(2)}(\underline{\theta})\right\}$ as the sample version of the least squares value with any $\underline{\theta}$. Denoting the gradient and Hessian by $\frac{\partial \hat{Z}_n(\underline{\theta})}{\partial\underline{\theta}}$ and $\frac{\partial^2 \hat{Z}_n(\underline{\theta})}{\partial\underline{\theta}^2}$, the mean value theorem gives $\frac{\partial \hat{Z}_n(\underline{\theta})}{\partial\underline{\theta}}-\frac{\partial \hat{Z}_n(\theta)}{\partial\underline{\theta}}=\frac{\partial^2 \hat{Z}_n(\check{\theta)}}{\partial\underline{\theta}^2}(\underline{\theta}_n-\theta)$ for some $\check{\theta}$ between $\theta$ and $\underline{\theta}$. Since $\frac{\partial \hat{Z}_n(\underline{\theta})}{\partial\underline{\theta}}=0$, the asymptotic distribution of the estimate $\underline{\theta_n}$ depends only on $\frac{\partial \hat{Z}_n(\theta)}{\partial\underline{\theta}}$ and $\frac{\partial^2 \hat{Z}_n(\check{\theta)}}{\partial\underline{\theta}^2}$. Applying the uniform law of large numbers to $\frac{\partial^2 \hat{Z}_n(\check{\theta)}}{\partial\underline{\theta}^2}$ and the CLT for orthogonal martingale differences to $\frac{\partial \hat{Z}_n(\theta)}{\partial\underline{\theta}}$ yields the asymptotic distribution in \ref{AsymptoticDistroFGARCH(1,1)estimator}. The full proof can be found in the Appendix.\\ My version

The proof of \Cref{AsymptoticDistroFGARCH11estimator} relies on a componentwise application of the mean value theorem. Define
\[
\hat{Z}_n(\underline{\theta})
=
\frac{1}{n}\sum_{i=2}^n
\left\{ y_i^{(2)}-\hat{s}_i^{(2)}(\underline{\theta})\right\}^\top
\left\{ y_i^{(2)}-\hat{s}_i^{(2)}(\underline{\theta})\right\},
\]
the sample least squares criterion for a generic parameter vector $\underline{\theta}$.
Denote its gradient and Hessian by
$\frac{\partial \hat{Z}_n(\underline{\theta})}{\partial\underline{\theta}}$
and
$\frac{\partial^2 \hat{Z}_n(\underline{\theta})}{\partial\underline{\theta}^2}$,
respectively.
Applying the mean value theorem componentwise to the gradient of $\hat{Z}_n$ between the true parameter $\theta$ and the estimator $\underline{\hat{\theta}}_n$ yields
\[
\frac{\partial \hat{Z}_n(\underline{\hat{\theta}}_n)}{\partial\underline{\theta}}
-
\frac{\partial \hat{Z}_n(\theta)}{\partial\underline{\theta}}
=
\frac{\partial^2 \hat{Z}_n(\check{\theta})}{\partial\underline{\theta}^2}
(\underline{\hat{\theta}}_n-\theta),
\]
for some intermediate value $\check{\theta}$ lying between $\theta$ and
$\underline{\hat{\theta}}_n$.
Since $\underline{\hat{\theta}}_n$ minimizes the criterion, its gradient vanishes, so that
\[
\sqrt{n}(\underline{\hat{\theta}}_n-\theta)
=
-
\left(\frac{\partial^2 \hat{Z}_n(\check{\theta})}{\partial\underline{\theta}^2}\right)^{-1}
\sqrt{n}\,
\frac{\partial \hat{Z}_n(\theta)}{\partial\underline{\theta}}.
\]
The asymptotic distribution of $\underline{\hat{\theta}}_n$ therefore follows from the limiting behavior of the gradient and the Hessian. More precisely, a uniform law of large numbers implies that the Hessian converges to $2\mathbf{Q}_0$, while a central limit theorem for orthogonal martingale differences yields the asymptotic distribution of the gradient evaluated at $\theta$. Combining these two results gives the asymptotic normality stated in \Cref{AsymptoticDistroFGARCH11estimator}. The full proof is deferred to the Appendix.\\
\citet{Cerovecki2019} proposes an alternative estimator based on pseudo quasi-maximum likelihood. Applying maximum likelihood techniques to functional data is generally not straightforward since a genuine likelihood for curves is typically not available. As a consequence, \citeauthor{Cerovecki2019} also assume that $\delta(t)$, $\alpha$, and $\beta$ can be decomposed into a finite combination of baseline functions similar to (\ref{basisdecompositiondelta}-\ref{basisdecompositionbeta}). Under this parametrisation, the projected squared process satisfies a recursion analogous to \Cref{FGARCH11MatrixForm}, which can be related to a semi-strong multivariate CCC-GARCH model with $R=\mathbb{I}_M$, where $\mathbb{I}_M$ denotes the $M\times M$ identity matrix.\\
\citet{Cerovecki2019} first establish consistency and asymptotic normality for the quasi-maximum likelihood estimator in this semi-strong multivariate CCC-GARCH setting and then use these results to derive consistency and asymptotic normality for the corresponding pseudo-QML estimator of the FGARCH model. The estimator is defined by
\begin{gather}
\label{QMLestimationequation}
    \underline{\hat{\theta}}_n= \arg \min_{\theta \in \Theta} \sum_{i=1}^n \tilde{\ell}_i(\theta), \quad \tilde{\ell}_i(\theta)=\sum_{m=1}^M\left\{\frac{\langle y_i^2,\varphi_m\rangle}{\langle \tilde{\sigma}_i^2,\varphi_m\rangle}+\log\langle \tilde{\sigma}_i^2,\varphi_m\rangle \right\},
\end{gather}
where the fitted volatility is calculated recursively as
\begin{gather}
    \tilde{\sigma}_i^2=\tilde{\sigma}_i^2(\theta) = \delta+\alpha y_{i-1}^2+ \beta \tilde{\sigma}_{i-1}^2, \quad \text{for }i=1,...,n.
\end{gather}
Finally, \citeauthor{Aue2017} discusses the choice of basis functions. In principle, any orthonormal system on $[0,1]$ can be used as basis functions. Common examples include B-splines, Fourier bases, and wavelet functions. As a data-driven alternative, one may estimate functional principal components (FPCs) from the squared observations and use these as basis functions. To do so, define the covariance kernel of $y_0^2(t)$ as
\begin{gather}
    D(t,s) = \text{Cov}(y_0^2(t),y_0^2(s)), \quad s,t\in[0,1].
\end{gather}
An estimate of $D(t,s)$ can then  be obtained by
\begin{gather}
    \hat{D}(t,s) = \frac{1}{n}\sum_{i=1}^n\{y_i^2(t)-\overline{y}_n^{(2)}(t)\}\{y_i^2(s)-\overline{y}_n^{(2)}(s)\}
\end{gather}
where $\overline{y}_n^{(2)}(t)=\frac{1}{n}\sum_{i=1}^ny_i^2(t)$. Using the spectral decomposition of $\hat{D}(t,s)$ the empirical eigenfunctions $\{\hat{\varphi}_1,...,\hat{\varphi}_M\}$ and eigenvalues $\{\hat{\lambda}_1,...,\hat{\lambda}_M\}$ are obtained. Note that under \labelcref{UnitVarianceepsilonfunction} $\mathbb{E}[y_i^2(t)|\mathcal{F}_{i-1}]=\sigma_i^2(t)$. Hence, the covariance structure of the squared observations contains information about the latent volatility process. This suggests using the leading empirical eigenfunctions of $\hat{D}$ as a data-driven basis, since they capture the dominant directions of variation in $y_i^2$ and therefore also in the unobservable volatilities. \citet{Aue2017} show that, if the empirical eigenfunctions are used in place of the deterministic basis $\Phi_M$ and $\int \mathbb{E}[y_0^4(t)]\,dt < \infty$ then the strong consistency result in \Cref{StrongConstistencyFGARCH11} remains valid. This extension is justified by the weak-dependence properties of the FGARCH$(1,1)$ process, in particular its geometric $L^2$-$m$-approximability established earlier.\\
The same basis functions cannot be used in this exact way when using the QML estimator since the quasi-maximum likelihood requires the basis functions to be non-negative. This ensures that the inner products in \labelcref{QMLestimationequation} are positive and thus the minimization problem is well defined and reaches its minimum on the compact set $\Theta$.\\
\citet{Cerovecki2019} therefore proposes to use families of non-negative and linearly independent functions, such as the power basis, the exponential basis, or polynomial bases that are non-negative on $[0,1]$. Alternatively, they develop a data-driven method based on Functional Principal Component Analysis (FPCA). Since standard principal components are orthonormal and therefore need not be non-negative, they cannot be used directly in the pseudo-QML framework. \citeauthor{Cerovecki2019} propose to shift the eigenfunctions obtained from FPCA using the following transformation. First, define $\phi_1(t)=1$ and $\phi_2(t)=\mu(t)$. Then, define the remaining basis functions by
\begin{gather}
    \phi_{m}(t) := \varphi_{m-2}(t)-\inf_{t\in[0,1]} \varphi_{m-2}(t) \wedge 0, \quad \text{for all } m=3,...,M
\end{gather}
where $\wedge$ denotes the minimum operator and $\inf_{t\in[0,1]} \varphi_{m-2}(t)$ is the minimum value of $\varphi_{m-2}$ over $[0,1]$. This transformation shifts each principal component upward, if necessary, so that the resulting function is non-negative on $[0,1]$. In practice, the mean function and eigenfunctions are replaced by their empirical counterparts. \citet{Cerovecki2019} report in their simulation study that this empirical, data-driven choice performs relatively well, even when compared with settings in which the true, but unobserved, basis functions from the data-generating process are used. \\
Using a set of linearly independent functions instead of orthonormal eigenfunctions comes at the cost of losing some of the structural advantages of an orthonormal basis. In particular, because of the loss of orthogonality, the basis representation does not retain the same computational simplicity and interpretability as in the case of FPCs. Moreover, the associated Gram matrix is, in general, no longer equal to the identity, which may complicate the finite-dimensional representation of the model and may lead to numerical instability if the basis functions are highly correlated. These issues have to be balanced against the non-negativity requirement of the quasi-maximum likelihood approach, which makes this choice of basis functions necessary. The practical relevance of this trade-off will be discussed further in \Cref{sec:Simulation}, where the comparison of the QML and LS estimators, in particular the results in \Cref{subsec:Simulation2}, may also reflect the effect of this choice of basis functions.
\subsection{Value-at-risk}
\label{subsec:ValueAtRisk}
This subsection provides a brief overview of the concept of VaR. After the historical overview given in the introduction, the present subsection turns to the formal definition of VaR, its main theoretical properties, and the central criticisms that have been raised against it.\\
Intuitively, VaR is intended to specify a threshold below which returns fall only with a predetermined probability. In probability terms, this means that, under the assumption that the map $x\mapsto \mathbb{P}(y_i\leq x\mid\mathcal{F}_{i-1})$ is strictly increasing, where $\mathcal{F}_{i-1}$ denotes the $\sigma$-algebra containing the information available up to time $i-1$, it holds that
\begin{gather}
    \mathbb{P}(y_i\leq \mathrm{VaR}_i^\tau \mid \mathcal{F}_{i-1})=\tau, \quad \tau\in(0,1).
\end{gather}
Here, $\mathrm{VaR}_i^\tau$ denotes the almost surely unique $\mathcal{F}_{i-1}$-measurable random variable corresponding to the quantile level $\tau$. In the intraday functional setting considered in this thesis, this definition is applied pointwise in $t$. More precisely, the intraday VaR curve is given by $\mathrm{VaR}_i^\tau(t)$, where for each $t\in[0,1]$ it holds that
\begin{gather}
    \mathbb{P}(y_i(t)\leq \mathrm{VaR}_i^\tau(t)\mid\mathcal{F}_{i-1})=\tau.
\end{gather}
Two important properties of VaR follow from this definition. First, VaR is not subadditive, meaning that the VaR of a combined portfolio can exceed the sum of the individual VaRs of the portfolios from which it is formed. As a consequence, VaR is not a coherent risk measure \citep{Artzner1999}. To see this in a simple scalar example, let $A$ and $B$ be two independent random variables with identical distributions given by
\begin{gather}
    \mathbb{P}(A=0)=0.95,\quad \mathbb{P}(A=-10)=0.05,
\end{gather}
and
\begin{gather}
    \mathbb{P}(B=0)=0.95,\quad \mathbb{P}(B=-10)=0.05.
\end{gather}
Then, at the confidence level $\tau=0.05$, the VaR of each individual random variable is equal to $0$. However, for the sum $A+B$, the probability of obtaining the value $0$ is
\begin{gather}
    \mathbb{P}(A=0 \text{ and } B=0)=0.95\cdot 0.95=0.9025.
\end{gather}
Hence, the lower $5\%$ quantile of $A+B$ is $-10$, so that the VaR of the combined position is $-10$.\\
A second important property of VaR is that it only captures the quantile at the chosen confidence level and therefore does not reflect the magnitude of losses beyond this threshold.\\
These two features, together with the widespread use of VaR as a risk management tool, have led to substantial criticism of VaR. As a result, extensions to VaR have been established. The most important extension to VaR is Expected Shortfall, also called Conditional VaR. Expected Shortfall in the intraday setting considered here is defined as follows:
\begin{gather}
    \mathrm{ES}_i^\tau(t)= \mathbb{E}[y_i(t)|\mathcal{F}_{i-1},y_i(t)<\mathrm{VaR}_i^\tau(t)].
\end{gather}
Expected Shortfall is subadditive and takes the behavior of the tail of the distribution into account. Therefore, Expected Shortfall addresses the two main theoretical issues of VaR. \citet{RICE20201023} also considers Expected Shortfall in the appendix of their paper. They show that the method explained in \Cref{sec:Simulation} can easily be extended to the concept of Expected Shortfall. This is beyond the scope of this thesis.
\section{Simulation Study}
\label{sec:Simulation}
%\begin{itemize}
%    \item Explain bootstrapping and Probability scoring measures
%\end{itemize}
In the next section, I will explore the finite sample properties of the FGARCH model and highlight a few practical considerations that must be made when using the FGARCH model for forecasting intra-day VaR in practice.\\
First, it should be emphasized that functional data are, in practice, observed on a finite grid. Consequently, both the properties discussed in \Cref{subsec:FGARCHModel} and the estimation procedure described in \Cref{subsec:FGARCHEstimation} are implemented in a discretized framework. This discretization does not yield a purely multivariate GARCH model; rather, it leads to a finite-dimensional approximation of the underlying functional model. In particular, operations defined in the functional setting—such as integrals in inner products—must be approximated numerically, which requires explicit consideration of the grid structure and the distances between adjacent grid points. For all simulations, functional data are approximated on an equidistant grid between $[0,1]$ with 100 grid points.\\
Second, to forecast VaR, an appropriate method has to be used to approximate the pointwise distribution of the innovation process $\varepsilon_i(t)$. Then the VaR forecast is equal to $\hat{\sigma}_i(t)\hat{\varepsilon}_i^\tau(t)$ where $\hat{\sigma}_i(t)$ is the forecasted value of $\sigma_i(t)$ using the fitted FGARCH model and $\hat{\varepsilon}_i^\tau(t)$ is the estimated $\tau$ quantile of $\varepsilon_i(t)$. To estimate $\hat{\varepsilon}_i^\tau(t)$ \citet{RICE20201023} propose three methods. For the first method, \citeauthor{RICE20201023} impose that $\varepsilon_i(t)$ follows a Ornstein-Uhlenbeck process. The second method estimates $\hat{\varepsilon}_i^\tau(t)$ using a bootstrap algorithm. Method three utilizes extreme value theory to better model the tails of the distribution. I will focus on the second method since it performed the best in practice \citep{RICE20201023}.\\
To obtain $\hat{\varepsilon}_i^\tau(t)$ using the bootstrap algorithm, first the fitted $\hat{\sigma}_i(t)$ needs to be computed. Next, $\hat{\varepsilon}_i(t)$ is calculated by $\hat{\varepsilon}_i(t)=\frac{y_i(t)}{\hat{\sigma}_i(t)}$. Finally, to get $\hat{\varepsilon}_i^\tau(t)$ the estimated error terms are resampled $B$ many times with replacement. $\hat{\varepsilon}_i^\tau(t)$ is then equal to the pointwise $\tau$-quantile of $(\hat{\varepsilon}_i^{(b)}(t))_{b\in[1,B]}$.\\
Third, to properly evaluate the results of the VaR forecasts I will employ the following statistical test. To test the accuracy of the intraday VaR forecast, I want to test whether the conditional coverage is at
the specified nominal level $\tau$. The $\mathcal{H}_0$ is thus
\begin{gather}
    \mathcal{H}_0 : \mathbb{E}[Z_i^\tau(t)-\tau]= 0 \text{ for all }t\in[0,1]
\end{gather}.
To test $\mathcal{H}_0$ so, define the pointwise violation process $Z_i^\tau(t)$ at quantile $\tau$
\begin{gather}
    Z_i^\tau(t) = \mathds{1}_{\{y_i(t)\leq \widehat{\mathrm{VaR}}_i^\tau(t)\}}; \quad i\in \{1,\dots,N\} \quad t\in[0,1]
\end{gather}
where $\mathds{1}$ denotes the indicator function and $\widehat{\mathrm{VaR}}_i^\tau(t)$ is the pointwise VaR forecast. \citet{RICE20201023} then proposes the following test statistic:
\begin{gather}
    T_n = N \|\overline{Z} - \tau \|^2.
\end{gather}
Here, $\overline{Z}$ denotes the sample mean $\overline{Z}=\frac{1}{N}\sum_{i=1}^NZ_i^\tau(t)$. \citet{RICE20201023} state that under $\mathcal{H}_0$ is distributed as follows
\begin{gather}
    T_n \overset{\mathcal{D}}{\to} \sum_{\ell=1}^\infty \lambda_\ell \mathcal{N}_\ell^2 \overset{\mathcal{D}}{\equiv} \Xi_1, \quad N \to \infty,
\end{gather}
where $(\mathcal{N}_\ell:\ell\geq 1)$ are i.i.d. standard normal distributed random variables, and $(\lambda_\ell:\ell\geq 1)$ are the eigenvalues of the covariance operator of the violation process, or more specifically, the eigenvalues of the kernel integral operator. The kernel of the integral operator is defined by $C_Z(t,s) = \mathrm{Cov}(Z_0(t),Z_0(s)), \quad t,s\in[0,1]$. One can then approximate the distribution of $\Xi_1$ via the use of Monte Carlo simulation after estimating $(\lambda_\ell:\ell\geq 1)$ from the sample version of the covariance operator. The p-value then is equal to $p=\mathbb{P}(T_n<\Xi_1)$.\\
Additional tests for evaluating the performance of the estimators and parameters chosen are introduced where needed.
\subsection{Simulation 1}
\label{subsec:Simulation1}
%\begin{itemize}
%    \item Similar to Aue et al. Show simulation where the method works.
%\end{itemize}
For the first simulation, a setup similar to \citet{Aue2017} and \citet{Cerovecki2019} is chosen. Set $\alpha(t,s)=\beta(t,s)=12t(1-t)s(1-s)$. \citet{Aue2017} and \citet{Cerovecki2019} both then define the error functions $\varepsilon_i(t)$ to follow a Ornstein-Uhlenbeck-type process. \citet{Cerovecki2019} then shows that under this setting, their (pseudo-) quasi-maximum likelihood estimator outperforms the least-squares estimator by \citet{Aue2017}. Ornstein-Uhlenbeck processes possess Gaussian increments that can be questioned for financial data that is known to have heavy tails. To test the results obtained in \citet{Cerovecki2019} using an alternative specification for the random innovation functions, I use the following specification for all $\varepsilon_i(t)$ 
\begin{align}
    \varepsilon_i(t) = e^{-\frac{t}{2}}T_i^\nu(e^t).
\end{align}
Here, $T_i^\nu$ denotes a Lévy process with increments with heavy tails approximating a Student-t distribution. $\nu$ denotes the typical degrees of freedom set for a standard Student-t distribution. The variance of the increments is scaled by $\frac{\nu-2}{\nu}$ to ensure that \labelcref{UnitVarianceepsilonfunction} holds. Additionally, I chose a different specification for $\delta$. To satisfy the assumptions made in (\ref{basisdecompositiondelta}-\ref{basisdecompositionbeta}) and the similar assumptions made for the QML estimator $\delta(t) =0.009\sqrt{30} t(1-t)$. In addition, I assume that the true basis function $\sqrt{30} t(1-t)$ is known. Under this setup, the optimal parameter vector is $\theta = \begin{pmatrix}
    d = 0.09 & a = 0.4 & b = 0.4
\end{pmatrix}$. The simulations were ran for three values of $\nu=\{5,10,25 \}$. In addition, three sample sizes were considered for the simulations to test the effect of larger samples.
\begin{table}[ht]
\label{TableMeanBiasSDLS}
\centering
\caption{Mean bias and standard deviation of estimators for LS}
\begin{tabular}{c ccc ccc ccc}
\toprule
\multirow{2}{*}{Sample size} & \multicolumn{3}{c}{$\nu = 5$} & \multicolumn{3}{c}{$\nu = 10$} & \multicolumn{3}{c}{$\nu = 25$} \\
\cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10}
& $d$ & $a$ & $b$ & $d$ & $a$ & $b$ & $d$ & $a$ & $b$ \\
\midrule
250 & \makecell[c]{.0077 \\[-2pt] (.0098)} & \makecell[c]{-.1104 \\[-2pt] (.1306)} & \makecell[c]{-.0632 \\[-2pt] (.2583)} & \makecell[c]{.0051 \\[-2pt] (.0072)} & \makecell[c]{-.0968 \\[-2pt] (.1420)} & \makecell[c]{-.0467 \\[-2pt] (.2229)} & \makecell[c]{.0055 \\[-2pt] (.0071)} & \makecell[c]{-.0934 \\[-2pt] (.1272)} & \makecell[c]{-.0404 \\[-2pt] (.2304)} \\
500 & \makecell[c]{.0042 \\[-2pt] (.0074)} & \makecell[c]{-.1000 \\[-2pt] (.1176)} & \makecell[c]{.0033 \\[-2pt] (.2162)} & \makecell[c]{.0033 \\[-2pt] (.0046)} & \makecell[c]{-.0869 \\[-2pt] (.1052)} & \makecell[c]{-.0088 \\[-2pt] (.1708)} & \makecell[c]{.0051 \\[-2pt] (.0066)} & \makecell[c]{-.0593 \\[-2pt] (.1126)} & \makecell[c]{-.0506 \\[-2pt] (.1921)} \\
1000 & \makecell[c]{.0042 \\[-2pt] (.0066)} & \makecell[c]{-.0904 \\[-2pt] (.1049)} & \makecell[c]{-.0069 \\[-2pt] (.1926)} & \makecell[c]{.0025 \\[-2pt] (.0040)} & \makecell[c]{-.0839 \\[-2pt] (.0817)} & \makecell[c]{.0260 \\[-2pt] (.1415)} & \makecell[c]{.0027 \\[-2pt] (.0044)} & \makecell[c]{-.0646 \\[-2pt] (.0877)} & \makecell[c]{.0001 \\[-2pt] (.1456)} \\
\bottomrule
\end{tabular}
\end{table}
\begin{table}[ht]
\label{TableMeanBiasSDQML}
\centering
\caption{Mean bias and standard deviation of estimators for QML}
\begin{tabular}{c ccc ccc ccc}
\toprule
\multirow{2}{*}{Sample size} & \multicolumn{3}{c}{$\nu = 5$} & \multicolumn{3}{c}{$\nu = 10$} & \multicolumn{3}{c}{$\nu = 25$} \\
\cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10}
& $d$ & $a$ & $b$ & $d$ & $a$ & $b$ & $d$ & $a$ & $b$ \\
\midrule
250 & \makecell[c]{.0026 \\[-2pt] (.0054)} & \makecell[c]{.0116 \\[-2pt] (.1487)} & \makecell[c]{-.0749 \\[-2pt] (.1934)} & \makecell[c]{.0017 \\[-2pt] (.0041)} & \makecell[c]{-.0150 \\[-2pt] (.1245)} & \makecell[c]{-.0472 \\[-2pt] (.1590)} & \makecell[c]{.0017 \\[-2pt] (.0041)} & \makecell[c]{-.0064 \\[-2pt] (.1106)} & \makecell[c]{-.0383 \\[-2pt] (.1478)} \\
500 & \makecell[c]{.0006 \\[-2pt] (.0033)} & \makecell[c]{-.0115 \\[-2pt] (.0954)} & \makecell[c]{-.0033 \\[-2pt] (.1284)} & \makecell[c]{.0008 \\[-2pt] (.0029)} & \makecell[c]{-.0078 \\[-2pt] (.0798)} & \makecell[c]{-.0231 \\[-2pt] (.1057)} & \makecell[c]{.0010 \\[-2pt] (.0025)} & \makecell[c]{.0201 \\[-2pt] (.0886)} & \makecell[c]{-.0396 \\[-2pt] (.0911)} \\
1000 & \makecell[c]{.0007 \\[-2pt] (.0024)} & \makecell[c]{.0084 \\[-2pt] (.0790)} & \makecell[c]{-.0201 \\[-2pt] (.0820)} & \makecell[c]{.0002 \\[-2pt] (.0017)} & \makecell[c]{-.0088 \\[-2pt] (.0586)} & \makecell[c]{.0051 \\[-2pt] (.0695)} & \makecell[c]{.0005 \\[-2pt] (.0017)} & \makecell[c]{.0016 \\[-2pt] (.0562)} & \makecell[c]{-.0159 \\[-2pt] (.0619)} \\
\bottomrule
\end{tabular}
\end{table}
From \Cref{TableMeanBiasSDLS} and \Cref{TableMeanBiasSDQML} two general observations can be inferred: First, a larger sample size and a specification for the random innovation functions with thinner tails leads to a more accurate estimation of the parameters. Second, QML-, generally, outperforms LS-estimation as was observed in \citet{Cerovecki2019}. This is, in particular, visible by the lower standard deviation while conclusions in terms of performance are more difficult to make on the basis of the mean bias. Although generally, the mean bias for the QML estimator seems to be lower in this simple setup.\\
Similar observations can be made when examining the p-values under this setup. QML also here seems to outperform LS estimation and thinner tails as well as a larger sample size leads to better estimates on average.
\subsection{Simulation 2}
\label{subsec:Simulation2}
%\begin{itemize}
    %\item Simulation 2 should demonstrate some of the weaknesses from the fGARCH model either by violating one of its assumptions or/and by giving reasons why the fGARCH model should be extended in some type of way (transition to fGARCH-X or fGARCH with multiple constant parameters or similar)
%\end{itemize}
In the second simulation, I consider a slightly more complex setup. First, the assumption made in (\ref{basisdecompositiondelta} - \ref{basisdecompositiondelta}) is violated by imposing a different specification for $\delta$, $\alpha$ and $\beta$. To do so, I keep $\beta(t,s)=12t(1-t)s(1-s)$, but change $\alpha$ to a Gaussian kernel defined as $\alpha(t,s) = 0.5 \exp(-(t-s)^2)$ and $\delta(t)$ to $\exp(-t^2)$.\\
Furthermore, to test the difference between the LS and QML estimator for $M>1$, I use two values for $M$, $M = \{3,5\}$. For the baseline functions, the respective data-driven method explained in \Cref{subsec:FGARCHEstimation} is used for both QML and LS estimation. This specifically also tests the difference between using orthonormal and solely linear independent baseline functions. As in \Cref{subsec:Simulation2} I use innovation functions with heavy tails and set the parameter vector for $\nu = \{5,25\}$. The sample size was set to $N=500$, and I ran $Q=50$ Monte Carlo simulations for each combination of the parameters. Due to the lack of computational resources only this small set of possible parameter combinations was considered.\\
Since testing performance for the fitted $\delta(t)$, $\alpha(t,s)$ and $\beta(t,s)$ cannot be done via the simple mean bias and standard deviation criterion used in \Cref{subsec:Simulation1} due to the differing set of baseline functions, an alternative metric has to be used. I follow the approach by \citet{Cerovecki2019} and define the relative mean squared deviations for $\delta$ and $\alpha$ by
\begin{gather}
    \frac{1}{Q^{1/2}\|\delta\|}
\left(
\sum_{q=1}^{Q} \|\hat{\delta}^{(q)} - \delta\|^2
\right)^{1/2},
\qquad
\frac{1}{Q^{1/2}\|\alpha\|}
\left(
\sum_{q=1}^{Q} \|\hat{\alpha}^{(q)} - \alpha\|^2
\right)^{1/2}.
\end{gather}
Here $\hat{\delta}^{(q)}$ denotes the estimated function for $\delta(t)$ for the $q$ simulation. $\| \|$ denotes the norm induced by the inner product as well as the operator norm respectively. The operator norm is approximated by the first eigenvalue of the discretized matrix form of $\hat{\alpha}^{(q)} - \alpha$. The relative mean squared deviations for $\beta$ are defined analogously.\\
One of the difficulties in using orthonormal instead of non-negative, linear independent basis function lies in the numerical optimization. If one estimates the parameter vector $\hat{\theta}_n = \begin{pmatrix}
    \mathbf{\hat{d}} & \mathbf{\hat{A}} & \mathbf{\hat{B}}
\end{pmatrix}$ as detailed in \Cref{subsec:FGARCHEstimation}, the resulting estimated discretized $\alpha$ and $\beta$ kernel matrices could end up not being pointwise strictly non-negative if the estimation is left unconstrained. However, since not $\hat{\theta}_n$ needs to be constrained, but the resulting $\alpha$ and $\beta$ kernels a specific solution is needed to handle this problem. As a consequence, I made use of the "Cobyla" algorithm \citep{CobylaAlgorithm} from the "NLopt" package \citep{NLopt}. The algorithm is a derivative-free optimization algorithm that allows the specification of equality and inequality constraints. A derivative-free optimization algorithm was chosen instead of an algorithm making use of the gradient since the gradient would have to be either approximated numerically or calculated explicitly which would involve a recursion over the entire sample size. I found that the gradient-free version performed equally well while being less computationally expensive as an algorithm that utilized the gradient.
\begin{table}[ht]
\centering
\caption{p-value for LS-estimation by number of principal components, parameter $\nu$, and quantile}
\begin{tabular}{c ccc ccc ccc}
\toprule
\multirow{2}{*}{Number of PCs}
& \multicolumn{3}{c}{$\nu = 5$}
& \multicolumn{3}{c}{$\nu = 10$}
& \multicolumn{3}{c}{$\nu = 25$} \\
\cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10}
& $q_{0.025}$ & $q_{0.01}$ & $q_{0.005}$
& $q_{0.025}$ & $q_{0.01}$ & $q_{0.005}$
& $q_{0.025}$ & $q_{0.01}$ & $q_{0.005}$ \\
\midrule
3 &  &  &  &  &  &  &  &  &  \\
5 &  &  &  &  &  &  &  &  &  \\
\bottomrule
\end{tabular}
\end{table}
\begin{table}[ht]
\centering
\caption{p-value for QML-estimation by number of principal components, parameter $\nu$, and quantile}
\begin{tabular}{c ccc ccc ccc}
\toprule
\multirow{2}{*}{Number of PCs}
& \multicolumn{3}{c}{$\nu = 5$}
& \multicolumn{3}{c}{$\nu = 10$}
& \multicolumn{3}{c}{$\nu = 25$} \\
\cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10}
& $q_{0.025}$ & $q_{0.01}$ & $q_{0.005}$
& $q_{0.025}$ & $q_{0.01}$ & $q_{0.005}$
& $q_{0.025}$ & $q_{0.01}$ & $q_{0.005}$ \\
\midrule
3 &  &  &  &  &  &  &  &  &  \\
5 &  &  &  &  &  &  &  &  &  \\
\bottomrule
\end{tabular}
\end{table}
Since functional data in practice can be represented in multivariate form it is natural to try to use a multivariate GARCH model to forecast intraday VaR. Due to the individual grid points of the data being highly correlated the PC-GARCH model could be an appropriate model for fitting such data. Consequently, I compared the finite sample performance of the PC-GARCH model under the current simulation specification to the results of the FGARCH$(1,1)$ model.
\begin{remark}{Comparison between the discretized FGARCH(1,1) model and the PC-GARCH model}{rem:FGARCHPCGARCH}
    As discussed previously, functional data are observed in practice on a finite grid and can therefore be represented in multivariate form. More precisely, a discretized version of the FGARCH$(1,1)$ model can be written as
    \begin{gather}
        y_i=\sigma_i \odot \varepsilon_i\\
        \sigma_i^2 = \delta+\alpha y_{i-1}^2+\beta\sigma_{i-1}^2
    \end{gather}
    where $\odot$ denotes the Hadamard product. Moreover, $y_i$, $\sigma_i$, $\varepsilon_i$, $\sigma_i^2$, $y_i^2$, and $\delta$ are vectors of length $T$, where $T$ is the number of grid points, while $\alpha$ and $\beta$ are $T \times T$ matrices. Here, $\sigma_i^2$ and $y_i^2$ denote the pointwise squares of $\sigma_i$ and $y_i$, that is, $\sigma_i^2 = \sigma_i \odot \sigma_i$ and $y_i^2 = y_i \odot y_i$.\\
    If, in addition, the basis functions in the finite-dimensional representation of the FGARCH model are estimated by FPCA, as was done for the LS estimator in \Cref{subsec:Simulation2}, then both the PC-GARCH and FGARCH models involve a reduction of dimensionality through a projection onto principal components. In this sense, the two approaches are similar in practice. Nevertheless, even under this specific setup, there remain important differences between them. First, while the PC-GARCH model models the full conditional covariance matrix of $y_i$ by assuming that it is spanned by a fixed set of principal components, the FGARCH model, under this finite-dimensional specification, is driven by a recursion for projected quantities associated with $y_i^2$ and $\sigma_i^2$, which are then combined with the basis functions to recover the pointwise conditional variance. Second, the principal components entering the PC-GARCH model are typically estimated from the covariance structure of $y_i$, whereas in the FGARCH framework they are obtained from the covariance structure of $y_i^2$. Third, while the FGARCH model allows dynamic interaction across components, the PC-GARCH model assumes that each latent variance component evolves through its own univariate GARCH recursion. Consequently, intuitively, the FGARCH model allows for cross-component dependence already at the level of the projected variance dynamics, while the PC-GARCH model keeps the latent variance dynamics separate and introduces dependence only through the reconstruction of the covariance matrix along fixed principal-component directions.
\end{remark}
\section{Discussion}
\label{sec:Discussion}
%\begin{itemize}
%    \item Summarize thesis
%    \item list possible extensions and literature that has already extended FGARCH
%    \item (Give outlook on future of risk measures? (Expected Shortfall?))
%\end{itemize}
% Eingehen auf die Research Fragen

In this thesis, the FGARCH model was presented, and sufficient conditions for a stationary solution were discussed. Furthermore, the difficulty of establishing necessary conditions for a stationary solution due to the non-equivalence of norms in the infinite-dimensional case was addressed. Following that, two estimation procedures for the FGARCH model were introduced and their finite sample performance was investigated using Monte Carlo Simulations. Both procedures rely on a dimension reduction assumption to a fixed finite dimension. While \citet{Cerovecki2019} discuss the extension to infinite-dimensional parameter spaces and prove consistency for the QML-estimator under this setup, the authors still acknowledge the difficulty of establishing an asymptotic distribution under this setting. The Monte Carlo simulations show that QML-estimation generally outperforms LS-estimation. Furthermore, I discuss the difficulties in practically implementing the LS estimator when using orthonormal eigenfunctions as basis functions.\\
% Limitations vielleicht hier einfügen
% Hier auch nochmal auf die Robustness der Results hinweisen bezogen auf die geringe Simulationsgröße
There are several directions in which the topic of this thesis can be extended. \citet{Kuehnert2020} proposes an alternative estimator for the FGARCH model based on yule-walker equations. The author also derives explicit asymptotic upper error bounds of estimation errors on finite-dimensional subspaces. His estimation procedure allows to estimate the functional operators directly without assuming a lower-dimensional representation as for the QML and LS estimator. Since his estimation approach would require a more detailed introduction, it was beyond the scope of this thesis.
Another direction in which the concept of FGARCH models can be further investigated is by proposing an alternative specification of the model. \cite{fTGARCH} proposes a functional threshold GARCH model that is meant to deal with the asymmetric response to the volatility process of unanticipated shocks. \citet{FGARCHX} introduces a FGARCH model with an additional exogenous covariate function that is meant to model long-range conditional heteroskedasticity. They use the model to forecast the intraday return curves of crude oil contracts.\\
\citet{FGARCHmultipleconst} establish a FGARCH model with multiple constant parameters and presents an estimation procedure for the model. The addition of multiple constant parameters extends the dimension of the space in which the conditional volatility curves are located in. They find that the FGARCH model with multiple constant parameters outperforms the standard FGARCH model in a real data application in both in-sample fitting and out-of-sample forecasting. Furthermore, \citet{aue2026operatorlevelarchmodel}, recently, introduced an operator-level ARCH model that is meant to model the full evolution of the conditional covariance operator rather than just the pointwise variance. One direction in which, to my knowledge, no research up to this date has been conducted is on incorporating additional functional processes into the FGARCH model to define a multivariate FGARCH model.\\ 
This thesis also does not address the issue of testing for conditional heteroskedasticity and order selection of an appropriate $(p,q)$ length of a FGARCH model. \citet{FHeteroTests} examines these issues and proposes statistical testing procedures for selecting an appropriate order and testing the assumption of conditional heteroskedasticity.\\
One further aspect in which this thesis could have been extended is the choice of basis functions used both for QML and LS estimation. In \citet{FGARCHX}, the authors of the FGARCH-X model discuss alternative methods for estimating non-negative baseline functions used for the QML-estimation of both the FGARCH model as well as their introduced functional GARCH-X model. They propose a sparse, non-negative version of FPCA as well as the use of truncated predictive factors based on the estimation procedure from \citet{KARGIN20082508}. Several other specifications of basis functions could be considered. One noteworthy alternative, to my knowledge yet to be explored in the context of the FGARCH model, are the smoothed FPC introduced by \citet{silverman1996smoothed}. The author shows that by replacing the usual $L^2$-orthonormality of the eigenfunctions by orthonormality with respect to an inner product that takes account of the roughness of the functions the estimation procedure effectively reduced the noise containted in the FPCs. This could be useful in the context of intraday return data which is known to display microstructure noise for a small enough sampling time frame \citep{Cerovecki2019}.\\
Finally, the estimation of value-at-risk for intraday return curves is also explored in other practical applications. \citet{kearney2025intradayfxvolatilitycurveforecasting} use the FGARCH model to forecast value-at-risk for intraday return curves of foreign exchange currency data for four major currencies. They also discuss two further specifications for basis functions.
\newpage
{\setstretch{0.6}
\bibliographystyle{agsm}
\bibliography{bibliography}
}

\appendix
\clearpage
\pagenumbering{Roman}
\setcounter{page}{1}
%\setcounter{tocdepth}{1}
\section{Appendix}
\label{sec:Appendix}
%\renewcommand{\thesubsection}{\Alph{subsection}}
\subsection{Availability of the Simulation Codes}
The simulation codes can be found in the following GitHub repository: \url{https://github.com/schmidtfabian/functional_garch}.
\subsection{Results of Section 2.1}
\subsubsection{Definition of Ergodicity}
This definition is taken from \citet{francq2019garch}.
\begin{definition}[Ergodic stationary process]
    A real valued strictly stationary process $(Z_i)_{i\in\mathbb{Z}}$ is said to be ergodic if and only if, for any Borel set $B$ and any integer $k$,
    \begin{gather}
        \frac{1}{n}\sum_{i=1}^n  \mathds{1}_B(Z_i,Z_{i+1},...,Z_{i+k})\rightarrow \mathbb{P}[(Z_1,...,Z_{1+k})\in B]
    \end{gather}
    with probability 1.
\end{definition}
It should be noted that the concept of ergodicity is much more general and can be extended to non-stationary sequences (see, for example, \citet{billingsley1995probability}).\\
The ergodic theorem applied to stationary sequences states the following.
\begin{theorem}[The ergodic theorem for stationary sequences]
    If $(Z_i)_{i\in\mathbb{Z}}$ is strictly stationary and ergodic, if $f$ is measurable and if 
    \begin{gather}
        \mathbb{E}|f(...,Z_{i-1},Z_i,Z_{i+1},...)|<\infty
    \end{gather}
    then
    \begin{gather}
        \frac{1}{n}\sum_{i=1}^n f(...,Z_{i-1},Z_i,Z_{i+1},...)\rightarrow\mathbb{E}[f(...,Z_{i-1},Z_i,Z_{i+1},...)]\quad a.s.
    \end{gather}
\end{theorem}
\subsubsection[Proof of Theorem 2.1]{Proof of \Cref{stationarityscalarGARCH11}}
As noted in \labelcref{RandomCoeffScalarGARCH} the conditional volatility equation of the scalar GARCH$(1,1)$ model can be rewritten into
\begin{gather}
    \sigma_t^2 = \delta + \gamma_{t-1} \sigma_{t-1}^2.
\end{gather}
Iterating yields
\begin{gather}
\label{IterationvolatilityscalarGARCH}
    \sigma_i^2 = \delta \left(1 + \sum_{j=1}^J \overbrace{\gamma_{i-1}\dots\gamma_{i-j}}^{\Gamma_{i,j}}\right)+\overbrace{\gamma_{i-1}\dots\gamma_{i-J-1}}^{\Gamma_{i,J}}\sigma_{i-J-1}^2\\
    := h_i(J) + \overbrace{\gamma_{i-1}\dots\gamma_{i-J-1}}^{\Gamma_{i,J}}\sigma_{i-J-1}^2
\end{gather}
Letting $J \rightarrow \infty$ identifies $h_t$ as a potential solution candidate
\begin{gather}
\label{StationarySolutionScalarGARCH}
    h_i := \delta \left(1+\sum_{j=1}^\infty \Gamma_{i,j}\right), \quad \text{where } \Gamma_{i-1} = \gamma_{i-1}\dots\gamma_{i-j}.
\end{gather}
To show that $h_i$ converges, I make use of the Cauchy Root Test for non-negative terms.
\begin{definition}[Cauchy Root Test]
    Let $\sum a_n$ be a series with non-negative terms and let $\lambda = \varlimsup a_n^{1/n}$. Then (i) if $\lambda<1$ the series $\sum a_n$ converges, (ii) if $\lambda>1$ the series $\sum a_n$ diverges.
\end{definition}
By the Cauchy Root Test, $\Gamma_{i,j}$ has to satisfy the following equation
\begin{gather}
    \lambda := \left[\Gamma_{i,j}\right]^{1/j} < 1.
\end{gather}
To show that this inequality holds, rewrite
\begin{gather}
    \lambda=\left[\Gamma_{i,J}\right]^{1/j} = \exp \left[\frac{1}{J}\,\log\left(\Gamma_{i,J}\right)\,\right] = \exp \left[\frac{1}{J}\,\log\left(\gamma_{i-1}\dots\gamma_{i-J}\right)\,\right]=\exp \left[\frac{1}{J}\,\sum_{j=1}^J\log\left(\gamma_{i-j}\right)\,\right].
\end{gather}
By an application of the law of large numbers $\frac{1}{J}\,\sum_{j=1}^J\log\left(\gamma_{i-j}\right)$ converges to $\mathbb{E}[ \log\{\overbrace{\alpha\varepsilon_t^2+\beta}^{\gamma_t}\}]$ and $\lambda<1$. Consequently, $h_i$ converges.\\
As a second step, we have to prove that $h_i$ is the unique solution to the scalar GARCH$(1,1)$ model. This proof is done by contradiction. To do so, assume that $\tilde{\sigma}_i^2$ denotes an alternative solution to the scalar GARCH$(1,1)$ model. By \labelcref{IterationvolatilityscalarGARCH}, $\tilde{\sigma}_i$ has to satisfy
\begin{gather}
    \sigma_i^2 = h_i(J) + \gamma_{i-1}\dots\gamma_{i-J-1}\tilde{\sigma}_{i-J-1}^2.
\end{gather}
For $\tilde{\sigma}_i$ to be an alternative solution to $h_i$ their difference has to be greater than $0$
\begin{gather}
    \tilde{\sigma}_i-h_i = (h_i(J)-h_i) + \gamma_{i-1}\dots\gamma_{i-J-1}\tilde{\sigma}_{i-J-1}^2.
\end{gather}
As $J\rightarrow\infty$, the right-hand side of the equation above goes to $0$. Thus, $h_i$ and $\tilde{\sigma}_i$ coincide. Therefore, $h_i$ is the unique solution of the scalar GARCH$(1,1)$ model.
That the solution is non-anticipative follows from \labelcref{StationarySolutionScalarGARCH}. Furthermore, \labelcref{StationarySolutionScalarGARCH} implies that there exists a functional $g$ such that
\begin{gather}
    \sigma_i^2=g(\varepsilon_{i-1},\varepsilon_{i-2},...), \quad i \in\mathbb{Z}
\end{gather}
which implies that $\sigma_i^2$ are Bernoulli shifts. The ergodicity of the solution is then implied by the work of \citet{stout1974almost}. \qed
\subsection{Proofs of Section 2.2}
\subsubsection[Proof of Theorem 2.2]{Proof of \Cref{StationarityFGARCH11}}
As stated previously, the proof of \Cref{StationarityFGARCH11} follows analogous arguments to the proof of \Cref{stationarityscalarGARCH11}. First, start by rewriting the volatility equation using the random coefficient form stated in \labelcref{FGARCH11volatilityalternative}
\begin{gather}
    \sigma_i^2(t) = \delta(t) + \int\gamma_{i-1}(t,s)\sigma_{i-1}^2(s)ds.
\end{gather}
Iterating backwards yields
\begin{gather}
    \sigma_i^2(t) = \delta(t) + \int\gamma_{i-1}(t,s)\sigma_{i-1}^2(s)ds\\ 
    = \delta(t) + \int\gamma_{i-1}(t,s_1)\left[\delta(s_1) + \int\gamma_{i-1}(s_1,s_2)\sigma_{i-2}^2(s_2)ds_2\right]ds_1\\
    = \delta_i(t) + \int \gamma_{i-1}(t,s_1)\delta_i(s_1)ds_1 + \int \int \gamma_{i-1}(t,s_1) \gamma_{i-2}(s_1,s_2)\sigma_{i-2}^2(s_2) ds_2ds_1\\
    = \delta_i(t) + \int \gamma_{i-1}(t,s_1)\delta_i(s_1)ds_1 + \int \int \gamma_{i-1}(t,s_1) \gamma_{i-2}(s_1,s_2)\delta(s_2) ds_2ds_1 +\\ \int \int \int \gamma_{i-1}(t,s_1) \gamma_{i-2}(s_1,s_2)\gamma_{i-3}(s_2,s_3)\sigma_{i-3}^2(s_3) ds_3ds_2ds_1
\end{gather}
and so on. Following that, define
\begin{gather}
    (\Gamma_{i,k}x)(t)=\int \cdots \int\gamma_{i-1}(t,s_1)\gamma_{i-2}(s_1,s_2)\cdots\gamma_{i-k}(s_{k-1},s_k)x(s_k)\,ds_k\cdots ds_2\,ds_1
\end{gather}
as in
\begin{align}
    (\Gamma_{i,1}x)(t)&=\int \gamma_{i-1}(t,s)x(s)ds \\
    \lVert\Gamma_{i,1}\rVert_{\mathcal{N}}&\leq \lVert\gamma_{i-1}\rVert_2
\end{align}
\subsection{Proofs of Section 2.3}
\begin{gather}
    \tilde{Z}_n(\underline{\theta})=\frac{1}{n}\sum_{i=2}^n\left\{ y_i^{(2)}-\tilde{s}_i^{(2)}(\underline{\theta})\right\}^\top \left\{ y_i^{(2)}-\tilde{s}_i^{(2)}(\underline{\theta})\right\} \rightarrow \sum_{m=1}^M\mathbb{E}\left[\langle\sigma_0^2(\varepsilon_0^2-1),\varphi_m \rangle^2 \right]+M(\underline{\theta}) \quad \text{a.s.}
\end{gather}
\subsection{Additional Tables and Plots from Section 3}
\begin{table}[ht]
\label{TablepvalueLS}
\centering
\caption{Mean p-value and standard deviation for LS}
\begin{tabular}{c ccc ccc ccc}
\toprule
\multirow{2}{*}{Sample size} & \multicolumn{3}{c}{$\nu = 5$} & \multicolumn{3}{c}{$\nu = 10$} & \multicolumn{3}{c}{$\nu = 25$} \\
\cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10}
& $q_{0.025}$ & $q_{0.01}$ & $q_{0.005}$ & $q_{0.025}$ & $q_{0.01}$ & $q_{0.005}$ & $q_{0.025}$ & $q_{0.01}$ & $q_{0.005}$ \\
\midrule
250 & \makecell[c]{.1682 \\[-2pt] (.0993)} & \makecell[c]{.0710 \\[-2pt] (.0390)} & \makecell[c]{.0524 \\[-2pt] (.0291)} & \makecell[c]{.1723 \\[-2pt] (.1957)} & \makecell[c]{.2841 \\[-2pt] (.1986)} & \makecell[c]{.2590 \\[-2pt] (.1278)} & \makecell[c]{.3425 \\[-2pt] (.2616)} & \makecell[c]{.1232 \\[-2pt] (.0524)} & \makecell[c]{.1602 \\[-2pt] (.0955)} \\
500 & \makecell[c]{.3096 \\[-2pt] (.3454)} & \makecell[c]{.4101 \\[-2pt] (.5436)} & \makecell[c]{.3514 \\[-2pt] (.4938)} & \makecell[c]{.1777 \\[-2pt] (.0634)} & \makecell[c]{.4422 \\[-2pt] (.3989)} & \makecell[c]{.4692 \\[-2pt] (.2856)} & \makecell[c]{.0472 \\[-2pt] (.0495)} & \makecell[c]{.0478 \\[-2pt] (.0281)} & \makecell[c]{.0036 \\[-2pt] (.0050)} \\
1000 & \makecell[c]{.6420 \\[-2pt] (.1201)} & \makecell[c]{.4400 \\[-2pt] (.2093)} & \makecell[c]{.3416 \\[-2pt] (.0844)} & \makecell[c]{.5624 \\[-2pt] (.0721)} & \makecell[c]{.4670 \\[-2pt] (.0538)} & \makecell[c]{.3946 \\[-2pt] (.0815)} & \makecell[c]{.8218 \\[-2pt] (.1220)} & \makecell[c]{.5034 \\[-2pt] (.2150)} & \makecell[c]{.3302 \\[-2pt] (.1537)} \\
\bottomrule
\end{tabular}
\end{table}
\begin{table}[ht]
\label{TablepvalueQML}
\centering
\caption{Mean p-value and standard deviation for QML}
\begin{tabular}{c ccc ccc ccc}
\toprule
\multirow{2}{*}{Sample size} & \multicolumn{3}{c}{$\nu = 5$} & \multicolumn{3}{c}{$\nu = 10$} & \multicolumn{3}{c}{$\nu = 25$} \\
\cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10}
& $q_{0.025}$ & $q_{0.01}$ & $q_{0.005}$ & $q_{0.025}$ & $q_{0.01}$ & $q_{0.005}$ & $q_{0.025}$ & $q_{0.01}$ & $q_{0.005}$ \\
\midrule
250 & \makecell[c]{.1037 \\[-2pt] (.0119)} & \makecell[c]{.0538 \\[-2pt] (.0033)} & \makecell[c]{.0413 \\[-2pt] (.0165)} & \makecell[c]{.1982 \\[-2pt] (.2378)} & \makecell[c]{.3790 \\[-2pt] (.1398)} & \makecell[c]{.3376 \\[-2pt] (.1720)} & \makecell[c]{.2686 \\[-2pt] (.0522)} & \makecell[c]{.1682 \\[-2pt] (.0464)} & \makecell[c]{.1404 \\[-2pt] (.0166)} \\
500 & \makecell[c]{.4475 \\[-2pt] (.3116)} & \makecell[c]{.5808 \\[-2pt] (.3321)} & \makecell[c]{.4636 \\[-2pt] (.3877)} & \makecell[c]{.1860 \\[-2pt] (.0329)} & \makecell[c]{.3405 \\[-2pt] (.2050)} & \makecell[c]{.4130 \\[-2pt] (.0781)} & \makecell[c]{.1688 \\[-2pt] (.2111)} & \makecell[c]{.0073 \\[-2pt] (.0070)} & \makecell[c]{.0156 \\[-2pt] (.0062)} \\
1000 & \makecell[c]{.5342 \\[-2pt] (.1316)} & \makecell[c]{.4191 \\[-2pt] (.2884)} & \makecell[c]{.3368 \\[-2pt] (.1129)} & \makecell[c]{.4516 \\[-2pt] (.0291)} & \makecell[c]{.6236 \\[-2pt] (.1406)} & \makecell[c]{.4518 \\[-2pt] (.1300)} & \makecell[c]{.7422 \\[-2pt] (.0697)} & \makecell[c]{.4950 \\[-2pt] (.2278)} & \makecell[c]{.3434 \\[-2pt] (.1781)} \\
\bottomrule
\end{tabular}
\end{table}
\clearpage
\section*{Declaration}
I hereby confirm that the work presented has been performed and interpreted solely by myself except for where I explicitly identified the contrary. I assure that this work has not been presented in any other form for the fulfillment of any other degree or qualification. Ideas taken from other works in letter and in spirit are identified in every single case.\\
\\
\\
\begin{minipage}{0.5\textwidth}
    Datum, Ort: 26.03.2026, Maintal
\end{minipage}
\begin{minipage}{0.5\textwidth}
    Unterschrift: \hrulefill
\end{minipage}
\end{document}
